### 2017-08-01 paper.tex
Add Evan's edits to Secs. I & II

\documentclass[aps,floatfix,prd,superscriptaddress,twocolumn]{revtex4-1}
% Intended to be built with pdflatex, using the REVTEX 4.1 package.
% Template taken from FermiLab APS Journal template.

% packages
\usepackage{amsmath}              % for \text, align, split ...
\usepackage{amssymb}              % for \gtrsim ...
\usepackage{braket}               % for \bra \ket
\usepackage{color}                % for colorful margin notes
\usepackage[mathscr]{euscript}    % for \mathscr
\usepackage{graphicx}             % for \includegraphics
\usepackage{hyperref}             % for \url and \href
\usepackage{listings}             % for typesetting code with \lstinline
\usepackage{tabularx}             % for tabularx

% commands/initialization
\hyphenation{ALPGEN} % avoid incorrect hyphenation
\hyphenation{EVTGEN} % ''
\hyphenation{PYTHIA} % ''
\lstset{basicstyle=\ttfamily} % define \lstinline for inline code snippet

\newcommand{\todo}[1]{\marginpar{\tiny{\textcolor{red}{#1}}}}
\renewcommand\todo[1]{} % if you wish to hide all todo notes

% *******************************************************************************
\begin{document}

\widetext
\leftline{Compiled \today}
\leftline{To be submitted to PRD}

\title{Elastic Scattering in General Relativistic Ray Tracing for Neutrinos}

\author{M.\ Brett Deaton}
\email{mbdeaton@ncsu.edu}
\affiliation{Joint Institute for Nuclear Astrophysics,
  Michigan State University, East Lansing, MI 48824, USA}
\affiliation{Department of Physics,
  North Carolina State University, Raleigh, NC 27695, USA}

\author{Evan O'Connor}
\affiliation{Department of Physics,
  North Carolina State University, Raleigh, NC 27695, USA}

\author{Y.\ L.\ Zhu}
\affiliation{Department of Physics,
  North Carolina State University, Raleigh, NC 27695, USA}

\author{Andy Bohn}
\affiliation{Center for Radiophysics and Space Research,
  Cornell University, Ithaca, NY 14853, USA}

\author{Jerred Jesse}
\affiliation{Department of Physics \& Astronomy,
  Washington State University, Pullman, WA 99164, USA}

\author{Francois Foucart}
\affiliation{Lawrence Berkeley National Laboratory,
  1 Cyclotron Rd., Berkeley, CA 94720, USA}

\author{Matthew D.\ Duez}
\affiliation{Department of Physics \& Astronomy,
  Washington State University, Pullman, WA 99164, USA}

\author{G.\ C.\ McLaughlin}
\affiliation{Department of Physics,
  North Carolina State University, Raleigh, NC 27695, USA}

% *******************************************************************************

\begin{abstract}
  We present a covariant ray tracing algorithm for computing high-resolution
  neutrino distributions in general numerical spacetimes with hydrodynamical
  sources.
  Our formulation treats the very important effect of
  elastic scattering of neutrinos off of nuclei and nucleons
  (changing the neutrino's direction but not energy)
  by incorporating estimates of the background neutrino fields.
  These background fields may be taken from a low-order moment simulation
  or be ignored, in which case the method
  reduces to a standard state-of-the-art ray tracing formulation.
  It handles radiation in regimes spanning optically thick to optically thin.
  We test the new code, highlight its strengths and weaknesses,
  and apply it to a simulation of a neutron star merger
  to compute neutrino fluxes and spectra,
  and to demonstrate a neutrino flavor oscillation calculation.
\end{abstract}

\maketitle

\section{Introduction}
Neutrinos are one of the dominant energy transport phenomena at play in
neutron star mergers, heating, cooling, and pushing the disrupted nuclear
matter.
In addition, they change the composition of the matter via charged current
interactions.
Because neutrinos scatter over length scales both large and small with
respect to fluid scales, accurate models require a neutrino treatment that
respects the freedom of neutrino distribution functions to vary
drastically from simple geometric distributions in thermodynamic equilibrium.

This is a challenging task in the environment of a merger,
which generally lacks any spatial symmetries, so that fully general
seven-dimensional solutions to the Boltzmann Equation are not feasible.
Leakage approximations
\citep{ruff1996-leakage_part1,ross2003-leakage_part2,
  deat2013-leakage, pere2016-asl, radi2016-dynamical}
\todo{cite others}
capture the qualitative effects of neutrinos on the matter, but
provide extremely limited information about the neutrino field itself.
\todo{discuss mc more clearly}
Monte Carlo methods like those
used in supernova simulations \cite{abdi2012-monte_carlo}
and stationary models of accretion disks \cite{rich2015-monte_carlo}
are an excellent tool but require large computational resources
because of their global treatment of the problem.

The state of the art today couples radiation and matter using a truncated
moment formalism
\citep{fouc2015-m1_nsbh, fouc2016-m1_nsns},
evolving the zeroth- and first-angular moments of the energy density,
that is the total energy- and momentum-densities.
This method, commonly called M1 transport, was formulated by
\cite{thor1981-truncated_moment} and modernized by
\cite{shib2011-truncated_moment}.
M1 transport has recently become popular in the core-collapse supernova
community as well
\cite{kuro2012-m1_ccsne,just2015-m1_code,ocon2015-gr1d_with_nu,
  skin2016-ray_vs_m1,robe2016-gr_m1_ccsne}.

For merger simulations \cite{fouc2016-m1_evolve_n} have recently
expanded their M1 transport code
to also evolve the zeroth-angular moment of the number density,
providing total energy densities and average energies throughout the
simulated volume.
Even so, with M1 transport codes only evolving two angular moments
and one or two energy moments,
they can extract only limited angular and spectral information
from the neutrino fields.

%\subsubsection*{What physics problems do we want to address?}
But many interesting unsolved problems require an accurate model of the neutrino
spectra and angular distributions.
With a model of the neutrino emission from a merger we can
1) examine neutrino effects on the nucleosynthesis of the ejected material
\citep{surm2011-nickel_56, robe2016-sph_nu_nucleo},
2) explore the rich flavor oscillation physics likely to occur
\citep{malk2012-mnr_1, malk2015-mnr_2, malk2016-mnr_3, zhu2016-mnr_nsns_remnant,
  vaan2016-uncovering_mnr},
3) improve closure relations used in M1 transport schemes
\citep{ramp2002-truncated_moment, shib2011-truncated_moment,
  card2013-truncated_moment, fouc2015-m1_nsbh, ocon2015-gr1d_with_nu}, and
4) study possible jet formation due to neutrino annihilation
\citep{ruff1999-nunubar_nsns, asan2000-nunubar, birk2007-nunubar,
  hari2010-gr_nunubar_collapsar, zala2011-nunubar, leng2014-nunubar}.

%\subsubsection*{Why ray tracing?}
In this work we present a ray tracing method to compute neutrino distribution
functions from state-of-the-art general relativistic M1 transport
hydrodynamics simulations.
We choose ray tracing because it is conceptually simple,
numerically inexpensive,
and extends to high resolution in energy and angle
by simply increasing the number of rays sampled.
Furthermore, the computational implementation parallellizes trivially.

With a ray tracing method we approach radiation transport from the perspective
of a single observer at a spacetime event $x_o^\alpha$.
Our goal is to compute the distribution function
$f^{\nu_\sigma}(x_o^\alpha;p_\beta)$, or the amount of neutrino radiation
of species $\nu_\sigma\in\{\nu_{\{e,\mu,\tau\}},\bar{\nu}_{\{e,\mu,\tau\}}\}$
with momentum $p_\beta$ impinging on $x_o^\alpha$.
To do so we trace a geodesic trajectory from $x_o^\alpha$ in the backwards
direction $-p_\beta$ to sample the incoming radiation along that line of sight.
By tracing a family of rays intersecting $x_o^\alpha$ we build up a
picture of the distribution function there.
And by sampling many observation points we construct a global picture
of $f^{\nu_\sigma}(x^\alpha;p_\beta)$.

The ray tracing framework
is conceptually simple because it solves the equation of radiation
transport (Eqn.~\ref{eqn:boltzmann}) along characteristics,
reducing it to a one-dimensional ordinary differential equation.
It is numerically cheap because it confines computations of
$f^{\nu_\sigma}(x_o^\alpha;p_\beta)$ to the past light-cone of $x_o^\alpha$,
with the history of that light-cone truncated at large optical depth.
\todo{state scaling $\mathscr{O}(n)$?}
And it easily extends to high resolution in energy and angle by simply
increasing the number of rays sampled.

%\subsubsection*{What is new and better about this ray tracing formulation?}
Several ray tracing formulations for radiation transport already exist.
Most formulations assume an analytic spacetime metric
\citep{birk2007-nunubar, caba2009-detecting_grb_nu,
  hari2010-gr_nunubar_collapsar, kova2011-gr_ray_tracing}.
And many make the simplifying assumption of blackbody emission from a
neutrinosurface
\citep{birk2007-nunubar, caba2009-detecting_grb_nu, kova2011-gr_ray_tracing},
limiting them to equilibrium, optically thick configurations.
Current state-of-the-art ray tracing formulations avoid the assumption of
blackbody spectra by integrating a local emissivity along each geodesic
(e.g. \cite{hari2010-gr_nunubar_collapsar} for neutrinos and
\cite{youn2012-gr_radiative_transfer} for photons).
But no formulations to date account for the important scattering and pair
processes outlined in Tab.~\ref{tab:neutrino_processes}.

We build upon existing ray tracing formulations
by eschewing any assumptions about the spacetime geometry,
integrating local emissivities,
and including elastic scattering in the integration along each geodesic.
To incorporate scattering processes in our method, we employ estimates of the
lowest-order moment of the neutrino distribution function
computed in an M1 transport simulation.
If moments are not available either from
an M1 transport evolution or a trustworthy analytical estimate,
our method reduces to the current state-of-the-art ray tracing method,
neglecting scattering.

%\subsubsection*{Why does a covariant formulation matter?}
We formulate our ray tracing equations covariantly---free from
assumptions about the spacetime geometry or coordinates. This is essential
because we want to apply the method as a postprocessing step using
time snapshots of data computed from general relativistic evolutions.
The spacetime represented in these snapshots is not analytic (i.e. Kerr).
And even in configurations that are described well by the Kerr metric
(e.g. a low-mass disk around a massive black hole),
the evolution coordinates are unlikely to present the metric in
a familiar analytic form.
This is because integrating the Einstein Equations often requires complicated,
time-dependent gauge conditions
\citep{lind2007-gen_harmonic, fouc2013-compactness_and_spin}.

%\subsubsection*{Why do scattering and pair processes matter?}
Elastic scattering (see Tab.~\ref{tab:neutrino_processes}) can signicantly
modify neutrino distributions in angle and dilute the emitted spectrum over a
larger emitting surface \citep{pere2016-asl}.
This is especially pronounced in the case of heavy-lepton neutrinos.
\todo{cite}
Inelastic scattering and pair processes can introduce further modifications.
Any phenomena that involve neutrino-neutrino interactions,
for example neutrino oscillations and neutrino-antineutrino annihilation,
\todo{cite}
depend sensitively on the angular distribution.
And spectral changes can strongly affect the nucleosynthesis
occuring in the ejected and irradiated material.
\todo{justify these statements}
The ray tracing method we present in this paper captures the dominant effects
of elastic scattering, while leaving the physics of inelastic scattering and
pair processes for later work.

%\subsubsection*{In what ways is this method limited?}
Ray tracing is ideally suited to problems requiring detailed knowledge of
radiation distribution functions over small regions of spacetime:
for example along a matter or radiation trajectory,
or over a small volume outside a source.

More fundamentally, though, ray tracing is limited by its naive
treatment of the Boltzmann Equation (Eqn.~\ref{eqn:boltzmann}),
a treatment which essentially decouples different neutrino momenta and species.
\todo{clarify how momenta are decoupled}
When we're solving for $f^{\nu_\sigma}(x^\alpha;p_\beta)$, we have no
information about $f^{\nu_\sigma}(x^\alpha;p'_\beta)$, and
$f^{\bar{\nu}_\sigma}(x^\alpha;p'_\beta)$, which are essential ingredients
for the source terms of the Boltzmann Equation describing the creation and
destruction of neutrinos due to the interaction processes described in
Tab.~\ref{tab:neutrino_processes}.

In this paper we outwit this limitation by incorporating coupled source terms
that depend on either previously-evolved or analytical
estimates of the neutrino fields.
We also retain the freedom to drop the coupling terms, in which case our method
becomes similar to existing state-of-the-art ray tracing methods.
Our method is not a standalone radiation transport scheme,
but serves as the final component of a hybrid scheme,
piggy-backing on a lower-order radiation transport method as a
post-processing step.

In Sec.~\ref{sec:formulation} we derive the ray tracing equations from the
Boltzmann Equation and describe our numerical scheme.
In Sec.~\ref{sec:tests} we present tests of the code.
In Sec.~\ref{sec:applications} we present an application
in the dynamical environment following the merger of two neutron stars,
computing neutrino flavor oscillation along an outgoing ray,
including the effects of coherent forward scattering with ambient neutrinos.

Greek tensor indices ($\alpha, \beta, ...$) range over all four coordinates,
whereas Latin indices ($i, j, ...$) range over the spatial coordinates 1--3.
We use naturalized units in which $\{\hbar,c,k_{\rm B}\}=1$.
And for most of the remainder of this article we suppress neutrino species label
$\nu_\sigma$ since the formulation is general to any species.
Where we do reference particular species we use the three categories
$\nu_e$, $\bar{\nu}_e$, and
$\nu_x=\left\{\nu_\mu,\bar{\nu}_\mu,\nu_\tau,\bar{\nu}_\tau\right\}$.

\section{Ray tracing formulation}
\label{sec:formulation}
The neutrino distribution function, $f(x^\alpha; p_\beta)$ is an invariant
quantity counting the number of neutrinos in a given six-volume of phase
space centered on $(x^\alpha,p_\beta)$.
The phase space volume elements are defined with respect to a fiducial
observer passing through event $x^\alpha$ with velocity $u^\alpha$:
\begin{align}
  \label{eqn:dV}
  dV & \equiv \sqrt{-\psi} \, dx \, dy \, dz \, u^t \\
  \label{eqn:dP}
  dP & \equiv \frac{1}{\sqrt{-\psi}} \, dp_x \, dp_y \, dp_z \,
  \frac{\varepsilon}{p^t},
\end{align}
where $\psi$ represents the determinant of the spacetime metric,
the index $t$ indicates the time-component of the given four-vector, and
\begin{equation}
  \label{eqn:varepsilon}
  \varepsilon \equiv -p_\mu u^\mu
\end{equation}
is the neutrino energy measured by our observer.
The number of particles in a given six-volume is
\begin{equation}
  dN=\frac{g}{(2\pi)^3}\,f\,dV\,dP,
\end{equation}
where $g$ counts the number of spin states accessible to the
particles ($g=1$ for neutrinos), and $f$ is the distribution function.
Each of $dV$, $dP$, $dN$, and $f$ are spacetime invariants
\citep{debb2009-gr_boltzmann_1, debb2009-gr_boltzmann_2, lind1966-gr_boltzmann}.

We may decompose the neutrino momentum like
\begin{equation}
  \label{eqn:def_momentum}
  p_\beta = \varepsilon (u_\beta + \ell_\beta),
\end{equation}
with $\ell_\beta$ the direction normal subject to the constraints
$u^\alpha \ell_\alpha = 0$ and $\ell^\alpha \ell_\alpha=1$.
With this decompositon we can write the arguments to the distribution function
$f(x^\alpha;\varepsilon,\ell_\beta)$.

Because $\ell_\beta$ is subject to two constraints
(normalization and orthogonality to the observer's velocity)
it has two degrees of freedom; we make this explicit by defining its
spatial cartesian components with respect to spherical polar angles
\begin{equation}
  \label{eqn:def_direction}
  \ell_\alpha \rightarrow
  q (s,\sin a \cos b,\sin a\sin b,\cos a),
\end{equation}
with $q$ and $s$ functions of $a$ and $b$.
Now our symbol for the distribution function,
$f(x^\alpha;\varepsilon,a,b)$,
makes manifest its seven independent arguments.

We also define a rotated frame
\begin{equation}
  \ell_{\alpha'}=\frac{\partial x^\alpha}{\partial x^{\alpha'}}\,\ell_\alpha \nonumber
\end{equation}
in which the cartesian components of the direction 1-form are defined
\begin{equation}
  \label{eqn:def_direction_primed}
  \ell_{\alpha'} \rightarrow
  q (s,\sin A \cos B,\sin A\sin B,\cos A),
\end{equation}
so that momenta with vanishing polar angle $A$ move outward along coordinate
radial rays.
This transformation is chosen so that for an observer far from the source,
incoming radiation will be concentrated into a narrow beam around $\cos A\approx1$,
independent of that observer's position in coordinate space.
For explicit definitions see App.~\ref{sec:def_momentum}.

In Minkowski spacetime, and for a stationary observer
$u_\beta\rightarrow(-1,0,0,0)$, we would have $s=0$, $q=1$,
and $\varepsilon=-p_t$. In that case $\cos A$ may be identified
\todo{cite someone that uses $\mu$}
with the familiar forward direction cosine $\mu$.

\subsection{Boltzmann Equation}
\label{ssec:boltzmann}
Neutrino radiation obeys the relativistic Boltzmann Equation,
which, suppressing the arguments of $f$ for simplicity, is written,
\todo{point out limit of QKEs}
\begin{equation}
  \label{eqn:boltzmann}
  \frac{d}{d\lambda}f = C[f],
\end{equation}
where $d/d\lambda$ denotes a derivative with respect to the affine
parameter defining the neutrino momentum (Eqn.~\ref{eqn:geodesic_x})
and $C[f]$ is the source term arising from interactions with the medium.
By Eqn.~\ref{eqn:geodesic_x},
the affine parameter has dimension ${\rm length}\,{\rm energy}^{-1}$,
so the source term has dimension ${\rm energy}\,{\rm length}^{-1}$.
The source term varies over phase space $(x^\alpha,p_\beta)$,
and depends locally on the distribution function $f$
and nonlocally on the distribution function of this neutrino and its
antiparticle at different momenta, $f'$ and $\bar{f}'$.
For simplicity we symbolize all of these dependencies with the shorthand $C[f]$.
The various neutrino interactions contributing to $C[f]$ are detailed in
App.~\ref{sec:source_terms}.

We make the right hand side of Eqn.~\ref{eqn:boltzmann} explicit by writing
the source term linear in $f$:
\begin{align}
  \label{eqn:boltzmann_linear}
  \frac{d}{d\lambda}f &=
  \mathscr{E} - \mathscr{K} f \\
  \label{eqn:boltzmann_linear_s}
  &= \mathscr{K}(\mathscr{S}-f)
\end{align}
where we have introduced
$\mathscr{E}$, the invariant total emissivity, and
$\mathscr{K}$, the invariant total opacity.
These describe respectively the energy gained and the energy lost
per length traveled by the neutrino, true scalar quantities
which take identical values for all observers.
In the second form we have introduced the source function
$\mathscr{S}\equiv\mathscr{E}/\mathscr{K}$, which makes manifest the
behavior of the right hand side, driving $f$ toward $\mathscr{S}$
over a lengthscale $\mathscr{K}^{-1}$ in our affine parameter;
or from Eqn.~\ref{eqn:proper_length}, over a proper lengthscale
$\varepsilon/\mathscr{K}$ measured by our fiducial observer.

These coefficients are computed by considering their dependence
on neutrino and antineutrino distribution functions at other momenta
(i.e.\ Fermi-blocking).
We consider the two dominant classes of interactions in this work:
the absorption/emission (AE) and elastic scattering (SE) processes
listed in Tab.~\ref{tab:neutrino_processes}.
Thus we separate these coefficients:
\begin{align}
  \label{eqn:inv_emissivity_components}
  \mathscr{E} &= \mathscr{E}_{\rm AE} + \mathscr{E}_{\rm SE},\\
  \label{eqn:inv_opacity_components}
  \mathscr{K} &= \mathscr{K}_{\rm AE} + \mathscr{K}_{\rm SE}.
\end{align}

The absorption/emission coefficients are computed from sums
over the relevant emissivities and opacities
for the reactions
\begin{align}
  \nu_e + n            &\leftrightarrow e^- + p,\nonumber\\
  \bar{\nu}_e + p      &\leftrightarrow e^+ + n,\nonumber\\
  \nu_e + {}^AZ        &\leftrightarrow e^- + {}^A(Z+1),\nonumber\\
  \bar{\nu}_e + {}^AZ  &\leftrightarrow e^+ + {}^A(Z-1),\nonumber
\end{align}
with ${}^ZA$ representing a nucleus of mass number $A$ and charge $Z$.
In terms of the emissivity $j(\varepsilon)$ describing number of neutrinos
of energy $\varepsilon$ emitted per length,
and the absorption opacity $\chi_a(\varepsilon)$
describing the number absorbed per length,
the coefficients are
\begin{align}
  \label{eqn:ae_emissivity_summed}
  \mathscr{E}_{\rm AE}(\varepsilon)
  &= \varepsilon \sum_{i\,{\rm reactions}} j_i(\varepsilon), \\
  \label{eqn:ae_opacity_summed}
  \mathscr{K}_{\rm AE}(\varepsilon)
  &= \frac{1}{1-f^{\rm eq}(\varepsilon)} \,
  \varepsilon \sum_{i\,{\rm reactions}} \chi_{{\rm a},i}(\varepsilon),
\end{align}
where $f^{\rm eq}$ is the distribution function of neutrinos in radiative
equilibrium with the matter, i.e. the Fermi-Dirac distribution function
\begin{equation}
  \label{eqn:feq}
  f^{\rm eq}(\varepsilon) \equiv
  \left(1+e^{\varepsilon/(k_{\rm B} T) -\eta_\nu}\right)^{-1},
\end{equation}
with the neutrino chemical potentials dependent on the local density,
temperature, and composition of the fluid via the neutron, proton,
and electron chemical potentials:
$\eta_{\nu_e}=-\eta_{\bar{\nu}_e}=\eta_p-\eta_n+\eta_{e^-}$ and
$\eta_{\nu_x}=0$.
The appearance of $f^{\rm eq}$ in Eqn.~\ref{eqn:ae_opacity_summed} is due to
the Fermionic nature of the neutrinos, causing $\mathscr{K}_{\rm AE}$ to be
different than the simple absorption opacity,
a phenomenon called stimulated absorption \cite{burr2006-neutrino_opacities}.
By detailed balance of the absorption/emission reactions, we may alternatively
write the emissivity in terms of the equilibrium distribution function:
\begin{equation}
  \label{eqn:ae_emissivity_feq}
  \mathscr{E}_{\rm AE}(\varepsilon)
  = \mathscr{K}_{\rm AE}(\varepsilon) f^{\rm eq}(\varepsilon).
\end{equation}
See App.~\ref{ssec:sources_ae} for details.
\todo{identify $\mathscr{K}_{\rm AE}$ and $\kappa^*$,
  in e.g.\ \cite{hari2010-gr_nunubar_collapsar}}

The elastic scattering coefficients are computed from
a background field, and a sum over opacities for
the reactions
\begin{align}
  \nu + N     &\leftrightarrow \nu + N,\nonumber\\
  \nu + {}^AZ &\leftrightarrow \nu + {}^AZ,\nonumber
\end{align}
with $N$ standing in for either $n$ or $p$.
In terms of a background field $\Phi(\varepsilon)$ describing
the number of neutrinos of energy $\varepsilon$ present at this event,
and the scattering opacity $\chi(\varepsilon)$ describing the
number scattered to other directions per length,
the coefficients are
\begin{align}
  \label{eqn:se_emissivity}
  \mathscr{E}_{\rm SE}(\varepsilon)
  &= \mathscr{K}_{\rm SE}(\varepsilon) \, \Phi(\varepsilon) \, \\
  \label{eqn:se_opacity_summed}
  \mathscr{K}_{\rm SE}(\varepsilon)
  &= \varepsilon \sum_{i\,{\rm reactions}} \chi_{{\rm s},i}(\varepsilon).
\end{align}
In the isotropic limit of trapped radiation,
$\Phi$ is equivalent to $f^{\rm eq}$;
in the free-streaming limit at a distance $r$ from a source,
$\Phi$ attenuates as $r^{-2}$.
See App.~\ref{ssec:sources_se} for details.

We may use Eqns.~\ref{eqn:ae_emissivity_feq} and \ref{eqn:se_emissivity}
to separate the source function thus
$\mathscr{S}=\mathscr{S}_{\rm AE}+\mathscr{S}_{\rm SE}$, with
\begin{align}
  \label{eqn:source_ae}
  \mathscr{S}_{\rm AE}
  &= \frac{\mathscr{K}_{\rm AE}}{\mathscr{K}} f^{\rm eq}, \\
  \label{eqn:source_se}
  \mathscr{S}_{\rm SE}
  &= \frac{\mathscr{K}_{\rm SE}}{\mathscr{K}} \Phi.
\end{align}
With these definitions we write a separated form of
Eqn.~\ref{eqn:boltzmann_linear_s}:
\begin{equation}
  \label{eqn:boltzmann_split}
  \frac{d}{d\lambda}f =
  \mathscr{K}_{\rm AE}(f^{\rm eq}-f)
  + \mathscr{K}_{\rm SE}(\Phi-f)
\end{equation}
From Eqn.~\ref{eqn:boltzmann_split} we see that
absorption/emission interactions drive the distribution function
toward $f^{\rm eq}$ over an affine lengthscale $\mathscr{K}_{\rm AE}^{-1}$,
and elastic scattering interactions drive it
toward $\Phi$ over an affine lengthscale $\mathscr{K}_{\rm SE}^{-1}$.
(Eqn.~\ref{eqn:proper_length} translates affine length to proper length
for a given observer; in this case the proper length scale is
$\varepsilon/\mathscr{K}$.)

This paper introduces the use of neutrino densities and fluxes evolved in an
M1 transport simulation to estimate the background field, $\Phi$.
App.~\ref{ssec:sources_se} details a method to calculate
$\Phi(\varepsilon)$ in two different ways:
\begin{itemize}
\item
  the \emph{spectral} method using densities and fluxes
  extracted from a simulation evolved over multiple energy groups
  to compute the background field with Eqn.~\ref{eqn:background_phi},
\item
  the \emph{gray} method using energy-integrated densities and fluxes
  extracted from a gray simulation and approximating the
  energy distribution with Eqns.~\ref{eqn:J_from_gray} and
  \ref{eqn:H_from_gray}.
\end{itemize}

\subsection{Trajectories}
\label{ssec:trajectories}
Each trajectory is uniquely labeled by a pair of vectors
giving some event on the trajectory, $x^\alpha$,
and the momentum at that event, $p_\beta$.
A family of trajectories shares the same $x^\alpha$; for example,
the family of emission trajectories is ($x^\alpha_e,p_\beta$),
and the family of observation trajectories is ($x^\alpha_o,p_\beta$).

Neutrino trajectories obey the geodesic equation, which may be decomposed into
the coupled first-order equations
\begin{equation}
\label{eqn:geodesic_x}
  \frac{d x^\alpha}{d\lambda} = p^\alpha,
\end{equation}
and
\begin{equation}
\label{eqn:geodesic_p}
  \frac{d p_\beta}{d\lambda} = -\Gamma^\alpha_{\beta\gamma} p^\gamma p_\alpha,
\end{equation}
where $p^\alpha=\psi^{\alpha\beta}p_\beta$,
$\psi^{\alpha\beta}$ is the inverse of the spacetime metric $\psi_{\alpha\beta}$,
and $\Gamma^\alpha_{\beta\gamma}$ are the standard connection coefficients,
\begin{equation}
  \label{eqn:christoffel}
  \Gamma^\alpha_{\beta\gamma} =
  \frac{1}{2} \psi^{\alpha\mu}
  (\psi_{\mu\beta,\gamma} + \psi_{\mu\gamma,\beta} - \psi_{\beta\gamma,\mu}),
\end{equation}
with the comma denoting a partial derivative
$\psi_{\mu\beta,\gamma}=\partial_\gamma\,\psi_{\mu\beta}$.

Each trajectory is parameterized by affine parameter, $\lambda$, increasing in
the direction of $\ell_\beta$. We label $\lambda=\lambda_e$ at $x^\alpha_e$,
as in Fig.~\ref{fig:affine_param}.
If we multiply Eqn.~\ref{eqn:geodesic_x} by $u_\alpha \equiv dx_\alpha/ds$,
we find the element of proper distance traversed by the
neutrino as measured by our fiducial obsever $u^\alpha$ is
\begin{equation}
  \label{eqn:proper_length}
  ds=\varepsilon \,d\lambda.
\end{equation}

\begin{figure}
  \includegraphics[width=\columnwidth]{fig-affine_parameter_label}
  % Script: $DIR/figures-radiation_transport_schematics.nb
  % w/ DIR=/Users/mbdeaton/Dropbox/Research/NuclearAstrophysics/NotesAndReference
  \caption{Affine parameterization of a neutrino trajectory of momentum
    $p_\alpha$.
    The fiducial observer with velocity $u^\alpha$ sits at $x^\alpha_o$,
    the neutrino emission event is at $x^\alpha_e$. The affine parameter
    increases from the emission event: $\lambda_e<\lambda'<\lambda''$.}
  \label{fig:affine_param}
\end{figure}

\subsection{The Formal Solution}
\label{ssec:rendering_eqn}
We can integrate Eqn.~\ref{eqn:boltzmann_split} directly, 
with our solution split into a boundary, absorption/emission, and
a scattering term, $f=f_{\rm bdry}+f_{\rm AE}+f_{\rm SE}$:
\begin{align}
  \label{eqn:rendering_fbdry}
  f_{\rm bdry}(\lambda,\lambda_e)
  &= f(\lambda_e) e^{-\tau(\lambda,\lambda_e)}, \\
  \label{eqn:rendering_fae}
  f_{\rm AE}(\lambda,\lambda_e)
  &= \int_{\lambda_e}^\lambda d\lambda'\,e^{-\tau(\lambda,\lambda')}
  \mathscr{K}_{\rm AE}(\lambda')f^{\rm eq}(\lambda'), \\
  \label{eqn:rendering_fse}
  f_{\rm SE}(\lambda,\lambda_e)
  &= \int_{\lambda_e}^\lambda d\lambda'\,e^{-\tau(\lambda,\lambda')}
  \mathscr{K}_{\rm SE}(\lambda')\Phi(\lambda'),
\end{align}
where the optical depth is defined,
\begin{equation}
  \label{eqn:optical_depth}
  \tau(\lambda,\lambda') \equiv \int_{\lambda'}^\lambda
  d\lambda'' \, \mathscr{K}(\lambda''),
\end{equation}
and the parameterization conventions are depicted in
Fig.~\ref{fig:affine_param}.
Note that Eqn.~\ref{eqn:optical_depth} employs the total absorption
plus scattering opacity, so that the optical depth attenuating the
integrands of Eqns.~\ref{eqn:rendering_fae} and \ref{eqn:rendering_fse}
is the total optical depth.

%Note the following equivalents; we could use any of these to express
%the integrands above:
%\begin{align}
%  \mathscr{K}_{\rm AE} \, f^{\rm eq}
%  &= \mathscr{K}\, \mathscr{S}_{\rm AE}
%  = \mathscr{E}_{\rm AE},\nonumber \\
%  \mathscr{K}_{\rm SE} \, \Phi
%  &= \mathscr{K} \, \mathscr{S}_{\rm SE}
%  = \mathscr{E}_{\rm SE}.\nonumber
%\end{align}

\subsection{Moments of the Distribution Function}
\label{ssec:moments}
\todo{code these up}
We may take angular moments of the distribution function:
\begin{align}
  \label{eqn:J}
  J(\varepsilon) &=
  \frac{\varepsilon^3}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta) \\
  \label{eqn:Ha}
  H^\mu(\varepsilon) &=
  \frac{\varepsilon^3}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta) \ell'^\mu \\
  \label{eqn:Sab}
  S^{\mu\gamma}(\varepsilon) &=
  \frac{\varepsilon^3}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta) \ell'^\mu \ell'^\gamma,
\end{align}
defining the specific energy density,
specific momentum density, and
specific radiation pressure tensor, respectively.
Here ``specific'' refers to the quantity being integrable over neutrino energy.
\todo{relate moments to other frames?}
Integrals~\ref{eqn:J}--\ref{eqn:Sab} are performed over a solid angle in
momentum space while holding $\varepsilon$ constant:
$d\Omega \equiv d(\cos a)\,db$.
\todo{consistent with Eqn.~\ref{eqn:def_direction}?}
We also make use of the number density and flux defined
\begin{align}
  \label{eqn:G}
  G(\varepsilon) &=
  \frac{\varepsilon^2}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta), \\
  \label{eqn:Ka}
  K^\mu(\varepsilon) &=
  \frac{\varepsilon^2}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta) \ell'^\mu.
\end{align}

The energy-integrated moments take the form
\begin{equation}
  \label{eqn:J_H_S_eps_integrated}
  X = \int_0^\infty d\varepsilon \, X(\varepsilon),
\end{equation}
with $X$ standing in for any of $\{J,H^\mu,S^{\mu\gamma}\}$ or
$\{G,K^\mu\}$,
the first having dimension ${\rm energy}\,{\rm length}^{-3}$,
and the second ${\rm length}^{-3}$.

There are two sensible observers to use to define the frame in which these
moments are computed, that is defining the four-velocity of
Eqn.~\ref{eqn:def_momentum}:
a stationary (also called an Eulerian) observer,
and a fluid (also called a comoving) observer \cite{smar1980-gr_hydro}.
Explicit definitions are given in App.~\ref{sec:def_momentum}.
We distinguish moments computed assuming an Eulerian observer with a tilde,
e.g. $\tilde{J}$, $\tilde{H}^\mu$, $\tilde{S}^{\mu\nu}$;
note that these three Eulerian moments are identical to the lab-frame moments
$E$, $F^\mu$, and $P^{\mu\nu}$
defined in
\cite{shib2011-truncated_moment, ocon2015-gr1d_with_nu, fouc2015-m1_nsbh}.

\subsection{Numerical Implementation}
\label{ssec:numerical}
Much of our numerical implementation is borrowed from the geodesic evolution
system described in~\cite{bohn2016-code}.
We integrate Eqns.~\ref{eqn:geodesic_x} and~\ref{eqn:geodesic_p} in the form
given by \citep{hugh1994-eh_finding},
and Eqns.~\ref{eqn:rendering_fae}, \ref{eqn:rendering_fse},
and \ref{eqn:optical_depth} in the form given below.
By using the time-component of Eqn.~\ref{eqn:geodesic_x} ($dt=d\lambda \, p^t$)
we may transform our integrations to coordinate time.
The coupled system of ordinary differential equations is
\begin{align}
  \label{eqn:num_x}
  \frac{dx^i}{dt} &=
  g^{ij} \frac{p_j}{p^t} - \beta^{i},\\
  \label{eqn:num_p}
  \frac{dp_i}{dt} &=
  -\alpha \alpha_{,i} p^t
  + \beta^k_{,i} p_k
  - \frac{1}{2} g^{jk}_{,i}\frac{p_j p_k}{p^t},\\
  \label{eqn:num_tau}
  \frac{d\tau}{dt} &=
  - \frac{1}{p^t}\mathscr{K},\\
  \label{eqn:num_fae}
  \frac{df_{\rm AE}}{dt} &=
  \frac{1}{p^t}e^{-\tau}\mathscr{K_{\rm AE}} \,f^{\rm eq},\\
  \label{eqn:num_fse}
  \frac{df_{\rm SE}}{dt} &=
  \frac{1}{p^t}e^{-\tau}\mathscr{K_{\rm SE}} \,\Phi.
\end{align}

We integrate each ray until we reach a terminal optical depth of
$\tau_{\rm term}$ at the earliest effective emission event $x_e^\alpha$.
The concept of earliest emission event is a facetious construct we use
to allow us to truncate our integration at an event along the ray where any
further additions to the field are negligible due to the large optical depth
between $x_e^\alpha$ and $x_o^\alpha$.
We choose $\tau_{\rm term}=14$ so that $e^{-\tau_{\rm term}}<10^{-6}$,
and we then discard the contribution of $f_{\rm bdry}$
(Eqn.~\ref{eqn:rendering_fbdry}).

Since we don't know the emission event a priori,
we follow the integration backwards in time,
from $t_o$ to $t_e$.
We begin each integration by setting initial values for the variables at $t_o$:
the observer specifies $x^i_o$ and $p_{i,o}$,
and we set $f_{{\rm AE},o}=f_{{\rm SE},o}=0$ and $\tau_o=0$.

We integrate Eqns.~\ref{eqn:num_x}--\ref{eqn:num_tau} with adaptive step sizes,
using the 3rd order Runge-Kutta algorithm
which produces an error estimate by comparing the 3rd and 2nd order solutions.
After each step is taken, the errors for each of the nine variables of
Eqns.~\ref{eqn:num_x}--\ref{eqn:num_tau} are compared.
If the maximum of these errors exceeds some absolute or relative threshold
\todo{specify ScaledAbsRel algorithm}
the step size is decreased and the step recomputed,
if it is below some threshold the next step size is increased.
% see OdeIntegrators::AdaptiveDense
% and OdeSteppers::RungeKutta3
% and OdeSteppers::DormandPrince5
% and OdeController
% and OdeControllers::ProportionalIntegral_Base::StepSizeToAttempt
% and OdeErrorMeasures::ScaledAbsRel
\todo{specify}

The volume data are components of the metric $\psi_{\alpha\beta}$,
its derivatives $\psi_{\alpha\beta,\gamma}$,
\todo{state num independent components}
and the hydrodynamical data $W$, $u_i$, $\varrho$, $T$, and $Y_e$.
These fields are stored in spectral representation.
If they are computed via hydrodynamical simulation, they are saved to disk
at specified coordinate times and interpolated
with spectral interpolation in space,
and 1st-order polynomial interpolation in time
(as described in \cite[App.~B]{bohn2016-code}).
If computed from a quasi-stationary solution to the
\todo{clarify quasistationary timescale}
general relativistic hydrodynamics equations,
a single time is stored to disk, and no time interpolation is needed.

\section{Code Tests}
\label{sec:tests}

To test our algorithm, we integrate the ray tracing equations
(Eqns.~\ref{eqn:num_x}--\ref{eqn:num_tau}) for various observers
in the following configurations. This suite of configurations
may be understood as a hierarchy of increasing physical realism:
beginning with a homogeneous medium of effectively
infinite extent (i.e.\ optically thick) and progressing to a model of a
1D pre-supernova, post-bounce collapse profile
evolved using an M1 transport hydrodynamical simulation.
In the pre-supernova model, we compare ray tracing distributions
to those calculated in a Monte Carlo transport simulation.
Finally, we calculate global measures of the neutrino fields outside of the
hypermassive neutron star and disk formed in a binary neutron star
merger simulated by \cite{fouc2016-m1_nsns, fouc2016-m1_evolve_n},
and compare ray tracing results to those from the M1 transport simulation.

In the following we present two forms of the integrated distribution functions,
Eqns.~\ref{eqn:num_fae} and \ref{eqn:num_fse}:
1) the full form  by which the solution is $f_{\rm scat}=f_{\rm AE,scat}+f_{\rm SE,scat}$
(labeled \emph{scat} since we include elastic scattering), and
2) the form treating only absorption/emission interactions
in which we set $\mathscr{K}_{\rm SE}=0$ so that $f_{\rm noscat}=f_{\rm AE,noscat}$
(labeled \emph{noscat} since we ignore elastic scattering).
There is no simple relation between these forms
(i.e.\ $f_{\rm AE,scat} \neq f_{\rm AE,noscat}$)
because $f_{\rm AE}$ is integrated using the absorption plus scattering optical depth
in the first form and solely the absorption optical depth in the second form.
The differences between the two methods is apparent
in Fig.~\ref{fig:cumulative_f_homogeneous}.
Before this work, the \emph{noscat} form of the equations has been the standard for
\todo{explain why; cite others}
ray tracing, though many authors include the scattering optical depth in
the integration of Eqn.~\ref{eqn:num_fae}
(see for example \cite{hari2010-gr_nunubar_collapsar}).

\subsection{Infinite Homogeneous Slab:
  Testing Thermodynamic Equilibrium}
\label{ssec:test_equilibrium}
In an optically thick region the radiation field is in thermodynamic equilibrium
with the matter.
We set up a large slab of matter
representative of the fluid in the Massive Collapsed Star test
(Sec.~\ref{ssec:test_collapse}) at the radius 50~km.
For comparision with that test we use the LS180 equation of state
\cite{latt1991-nuc_eos}
\footnote{available at \lstinline{stellarcollapse.org}}
in which the equilibrium $\nu_e$ neutrino chemical potential is
$\eta_{\nu_e}=-0.1555$
(with $\eta_{\bar{\nu}_e}=-\eta_{\nu_e}$ and $\eta_{\nu_x}=0$).
The opacity table is computed using \lstinline{NuLib} from cross-sections
neglecting weak magnetism
\todo{anything else?}
and is stored on a grid covering energy, density, temperature, and composition
ranges spanning $\varepsilon_i\in[1,280.5]$~MeV logarithmically,
$\varrho\in[10^6,6.3\times10^{15}]\,{\rm g}\,{\rm cm}^{-3}$ logarithmically,
$T\in[0.05,200]$~MeV logarithmically, and 
$Y_e\in[0.035,0.55]$ linearly,
with grid extents$\{18,82,65,51\}$ respectively.
Interpolation is performed linearly.

We use the equilibrium distribution functions to define background
fields for the \emph{scat} case:
in the \emph{spectral} method we use
$J(\varepsilon)=\varepsilon^3 f^{\rm eq}(\eta_\nu,T;\varepsilon)/(2\pi^2)$,
and in the \emph{gray} method we use
$J=C T^4 \mathscr{F}_3(\eta_\nu)/(2\pi^2)$ and
$G=C T^3 \mathscr{F}_2(\eta_\nu)/(2\pi^2)$,
using the Fermi integrals defined in Eqn.~\ref{eqn:fermi_integral},
and with $C$ the conversion constant from ${\rm energy}^3$ to ${\rm length}^{-3}$.

Fig.~\ref{fig:mfps-50km} presents the neutrino mean free paths
at this thermodynamic point over two energy decades.
For this test we set up our slab large enough for neutrinos of all
energies to be trapped.
Our domain extends to $\tau_{\rm term}\,L_{\rm mfp,max}\approx10^7\,{\rm km}$,
since our ray tracing algorithm integrates rays to terminal optical
depths of $\tau_{\rm term}\geq14$.

\begin{figure}
  \includegraphics[width=\columnwidth]{figure-mfps-50km}
  % Data:   $DIR/table-50km-nu?.dat
  % Script: $DIR/GnuplotScripts/plot-mfp-50km.gnu
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/HomogeneousIsotropic
  \caption{Mean free paths
    representative of the collapse profile at
    50~km ($\varrho=10^{11}\,{\rm g}\,{\rm cm}^{-3}$,
    $T=3.7$~MeV, $Y_e=0.12$).
    In the fluid rest frame these are given by
    $(\chi^*_a)^{-1}$ for the absorption
    and $(\chi_s)^{-1}$ for elastic scattering interactions.
    Note that elastic scattering opacities are identical across species
    below energies at which weak magnetism plays a role \cite{horo2002-weak_mag}.
    In this data they are exactly identical
    because we have turned off weak magnetism in our opacity calculations
    in order to compare our results with the historical literature.
    Computed using \lstinline{NuLib}.
  }
  \label{fig:mfps-50km}
\end{figure}

Fig.~\ref{fig:cumulative_f_homogeneous} displays the cumulative distribution
function integrated along the ray, for each of the three species and using
both methods, \emph{noscat} and \emph{scat}. We display
Eqns.~\ref{eqn:num_fae} and \ref{eqn:num_fse} in their integral form:
\begin{equation}
  \label{eqn:cumulative_f}
  f(t)
  = - \int_{0}^t \frac{dt'}{p^t}e^{-\tau(0,t')}
  \mathscr{K}(t')f^{\rm eq}(t'),
\end{equation}
and the integration proceeds backwards in time, from $t=0$
\todo{could write more simply because Minkowski}
to some terminal $t<0$.
The figure shows this backwards-in-time integration proceeding left to right.

\begin{figure}
  \includegraphics[width=\columnwidth]{fig-cumulative_f-homogeneous}
  % Data:   $DIR/Lev1-v0-ShootANeutrino-01/Lev1
  % Script: $DIR/GnuplotScripts/plot_f-integral_over_t.gnu
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/HomogeneousIsotropic
  \caption{Cumulative distribution functions at $\varepsilon=11.25$~MeV
    in homogeneous infinite slab of Sec.~\ref{ssec:test_equilibrium}.
    The integration proceeds from left to right, or backwards in time $t$.
    Each ray terminates when it achieves a total optical depth
    greater than $\tau_{\rm term}=14$.
    The positions of the points (only plotted for the $f_{\rm AE}$ components)
    are the time steps chosen by the adaptive time-stepping algorithm
    described in Sec.~\ref{ssec:numerical}.
  }
  \label{fig:cumulative_f_homogeneous}
\end{figure}

In Fig.~\ref{fig:cumulative_f_homogeneous} we see some expected features.
The final distribution functions asymptote to their equilibrium Fermi-Dirac
levels $f^{\rm eq}_{\{\nu_e,\bar{\nu}_e,\nu_x\}}=\{0.039, 0.053, 0.046\}$
within a few mean free paths;
in the \emph{scat} cases it is the sums $f_{\rm AE}+f_{\rm SE}$ that
achieve these values.
And the lengths of the rays are proportional to the mean free paths:
$L_{\nu_x}>L_{\bar{\nu}_e}>L_{\nu_e}$;
in the \emph{scat} cases these lengths are less than in the \emph{noscat} cases,
since the total mean free paths are less than the absorption mean free paths
(significantly so for $\nu_x$, negligibly so for $\nu_e$).

As Fig.~\ref{fig:mfps-50km} shows, integrating the distribution functions
over the energies $\varepsilon\in[0,100]$~MeV
probes our numerical solution over length scales from
0.1~km to $10^{5}\,{\rm km}$.
Fig.~\ref{fig:homogeneous_isotropic} shows the error in our results for
the \emph{noscat} case.
In this simple case, the dominant source of error is the neglected
boundary term with a relative scale of $e^{-\tau_{\rm term}}\approx10^{-6}$.
In the following inhomogeneous configurations,
errors from the integration dominate over the boundary term.
In this test and only this test, we used a higher-order integrator, a 5th
order Dormand-Prince algorithm \cite{pres2007-nr_3rd_ed}. This is because
\todo{discuss with Matt}
our default 3rd order Runge-Kutta algorithm estimates a vanishing error
in this configuration.

\begin{figure}
  \includegraphics[width=\columnwidth]{figure-ferr-50km}
  % Data:   $DIR/Lev1-v0-noscat-01/Lev1
  % Script: $DIR/GnuplotScripts/plot_ferr.gnu
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/HomogeneousIsotropic
  \caption{Relative error in integrated equilibrium distribution functions
    in the infinite homogeneous slab  described in
    Sec.~\ref{ssec:test_equilibrium}.
    Presented here are the relative differences between the final
    $f_{\rm AE}$ in the \emph{noscat} case
    (see Fig.~\ref{fig:cumulative_f_homogeneous})
    and the equilibrium distribution functions given by
    Eqn.~\ref{eqn:feq}.
    The source of this error is the discarded boundary term
    $f_{\rm bdry}\sim10^{-6}$, described in Sec.~\ref{ssec:numerical}.
  }
  \label{fig:homogeneous_isotropic}
\end{figure}

\subsection{Infinite Homogeneous Moving Slab:
  Testing Doppler Shift}
\label{ssec:test_doppler}
We reproduce the test above, but with the matter and observer in relative
motion.
We use a stationary (i.e.\ Eulerian) observer
and a fluid velocity $u^\alpha \rightarrow W(1,0,0,v)$ with $W=(1-v^2)^{-1/2}$.
All other thermodynamic variables and background fields are unchanged
since our ray integration uses these quantities in the fluid frame.

A stationary observer sees a neutrino with momentum
$p_\alpha \rightarrow \tilde{\varepsilon}(-1,\Omega_i)$,
with direction $\Omega_i \rightarrow (\sin A, 0, \cos A)$,
and observed energy $\tilde{\varepsilon}$.
In the fluid frame this neutrino has energy $\varepsilon=-u^\alpha p_\alpha$;
therefore the average energy varies with observing angle:
\begin{equation}
  \label{eqn:doppler_avg_eps}
  \langle \tilde{\varepsilon} \rangle(\cos A) =
  \langle \varepsilon \rangle^{\rm eq} \frac{1}{W(1-v \cos A)}.
\end{equation}
The equilibrium average energy $\langle \varepsilon \rangle^{\rm eq}$
is given by $T \mathscr{F}_3(\eta_\nu)/\mathscr{F}_2(\eta_\nu)\approx 3T$,
with the Fermi integrals given in Eqn.~\ref{eqn:fermi_integral}.
Eqn.~\ref{eqn:doppler_avg_eps} describes the well-known Doppler effect.

\begin{figure}
  \includegraphics[width=\columnwidth]{fig-moving_fluid_avg_eps}
  % Data:   $DIR/Lev1-v*-noscat-01
  % Script: $DIR/plot-avg_eps_vs_cosA.gnu
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/MovingFluid
  \caption{Average energy of $\bar{\nu}_e$ neutrino fields
    measured by a stationary observer,
    $\langle\tilde{\varepsilon}\rangle(\cos A)$.
    The observer sees neutrinos moving with the fluid for angles $\cos A<0$.
    Test described in Sec.~\ref{ssec:test_doppler}.
    The points are computed from ray tracing spectra; the lines are the
    analytic formula Eqn.~\ref{eqn:doppler_avg_eps}.
  }
  \label{fig:avg_eps_doppler}
\end{figure}

We sample the distribution function $f(\tilde{\varepsilon},\cos A,B)$
with ray tracing
over a uniform grid of 40 points in energy $\tilde{\varepsilon}\in[0,100]$
and 30 points in angle $\cos A\in[-1,1]$, holding fixed $B=\pi$.
Results are shown in Fig.~\ref{fig:avg_eps_doppler} for
the velocities $v=\{0,0.1,0.8\}$.
The ray tracing results are computed from total densities in each angular bin,
that is
\begin{equation}
  \label{eqn:avg_eps_per_cosA}
  \langle \tilde{\varepsilon} \rangle (\cos A) =
  \frac{\tilde{J}(\cos A)}{\tilde{G}(\cos A)},
\end{equation}
with the Eulerian energy density per angular bin given by
\begin{equation}
  \label{eqn:J_per_cosA}
  \tilde{J}(\cos A) =
  \frac{1}{(2\pi)^3} \int_0^{\infty} d\tilde{\varepsilon} \, \tilde{\varepsilon}^3
  \int_0^{2\pi} dB \, f(\tilde{\varepsilon},\cos A,B),
\end{equation}
and $\tilde{G}(\cos A)$ defined analogously.
\todo{add metric terms for these integrals}
Results are identical for all \emph{scat} and \emph{noscat} methods.
In Fig.~\ref{fig:avg_eps_doppler}
we see the common features of a red-shifted spectrum for receding fluid
($\cos A\ll0$),
a blue-shifted spectrum for approaching fluid
($\cos A\gg0$),
and a slightly red-shifted spectrum for fluid moving transverse to the
observer ($\cos A=0$).

\subsection{Idealized Star:
  Testing the Decoupling Regime}
\label{ssec:test_ab_star}
The homogeneous configurations of the previous sections may be extended to
probe the solution outside the optically thick regime
by setting up an idealized homogeneous star, with the thermodynamic variables
constant inside radius $R$ and vanishing outside.
We choose $R=50$~km and place our observer at $r=75$~km,
at which position there is a radiation cone of half-opening angle
$\cos A_{\rm max}\approx0.75$.

The formal solutions of
Eqns.~\ref{eqn:rendering_fae} and \ref{eqn:rendering_fse}
may be directly integrated in this scenario. Assuming Minkowski spacetime
and stationary fluid we have
\begin{align}
  \label{eqn:fae_homogeneous_general}
  f_{\rm AE} &=
  \frac{\chi_a^{*}}{\chi} f^{\rm eq} \left(1-e^{-\chi s}\right) \\
  \label{eqn:fse_homogeneous_general}
  f_{\rm SE} &=
  \frac{\chi_s}{\chi} \Phi \left(1-e^{-\chi s}\right) \\
  \label{eqn:geometric_s}
  s &= 2 R \left(1-\frac{r^2}{R^2}(1-\cos^2 A)\right)^{1/2},
\end{align}
where $s$ is the path length traversed by the ray through the star.

Since no analytic form is known for the background field $\Phi$,
we examine only the \emph{noscat} case,
with $\chi_s=0$ and $\chi=\chi_a^{*}$.
This scenario has been widely used in the literature as a test for
radiation codes
\cite{smit1997-two_moment,abdi2012-monte_carlo,fouc2015-m1_nsbh}.
We sample the distribution function over a uniform grid of
30 points in angle $\cos A \in [0.734,1]$ and
40 points in energy $\varepsilon \in [0,100]$~MeV.
Because of the step function in fluid variables at radius $R$,
we throttle the maximum step size at 1.25~km.

In Fig.~\ref{fig:f_absorption_sphere} we display the samples at
$\varepsilon=11.25$~MeV, along with the analytic functions
specific to each species' equilibrium distribution function and opacity.
As expected only $\nu_e$ saturates at $f^{\rm eq}$, remaining almost
constant across $\cos A$ until we get to rays that pass through a length
of the star comparable to or less than the mean free path at this energy,
16~km. We also see that $\bar{\nu}_e$ comes close to saturating with
a mean free path just over 100~km; and $\nu_x$ is well into the optically
thin regime.

Expanding the exponential in Eqn.~\ref{eqn:fae_homogeneous_general},
we see the distribution function takes the limiting values:
\begin{equation}
  \label{eqn:fae_homogeneous_limits}
  f_{\rm AE}(\cos A) = f^{\rm eq}
  \begin{cases}
    \chi s(\cos A) & s \ll \chi^{-1}\,\textrm{optically thin}\\
    1                  & s \gg \chi^{-1}\,\textrm{optically thick}.
  \end{cases}
\end{equation}
This agrees with Fig.~\ref{fig:f_absorption_sphere}, with $\nu_x$
in the optically thin limit and $\nu_e$ in the optically thick limit.

\begin{figure}
  \includegraphics[width=\columnwidth]{fig-f-absorption_star-11MeV}
  % Data:   $DIR/Lev1-01/Lev1
  % Script: $DIR/Lev1-01/Lev1/plot_f_monochromes-11.25MeV.gnu
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/HomogeneousAbsorptionSphere
  \caption{Distribution functions outside an idealized homogeneous star
    with radius $R=50$~km, and observer at $r=75$~km,
    described in Sec.~\ref{ssec:test_ab_star}.
    $A$ is the angle between the neutrino momentum and the $\hat{r}$ direction.
    In this plot we display only the samples at energy
    $\varepsilon=11.25\,{\rm MeV}$.
    The points are computed from ray tracing; the lines from the analytic
    solution, Eqn.~\ref{eqn:fae_homogeneous_general}.
  }
  \label{fig:f_absorption_sphere}
\end{figure}

Though we do not display it here,
we also examine $\langle \varepsilon \rangle(\cos A)$
as in Eqn.~\ref{eqn:avg_eps_per_cosA}, and robustly reproduce the
common feature of increasing average energies at the limb of the star
for $\nu_e$ and $\bar{\nu}_e$. In contrast $\langle \varepsilon \rangle$
for $\nu_x$ is constant across angles, as expected when the star
is optically thin at even the highest energies $\varepsilon\sim100$~MeV.

\subsection{Idealized Compact Star:
  Testing Gravitational Redshift and Geodesic Curvature}
\label{ssec:test_gravity}
Optically thick star in Schwarszchild.
Total luminosity at infinite separation may be computed by
computing the radial energy flux at a finite separation from the star, $r$,
and using $p_t$ for the energy. We compute this asymptotic luminosity
at several radii, $r$, and show that the luminosity is independent of
the observer's radius.
\todo{compute $L$, put in figure}

\subsection{Post-Bounce Collapse Profile:
  Testing Scattering}
\label{ssec:test_collapse}
To test our scattering treatments we calculate neutrino fields outside a
collapsed 15~${\rm M}_\odot$ star, 100~ms after core bounce,
comparing ray tracing fields to those from a Monte Carlo transport calculation.
Elastic scattering interior to the shock at $r\approx150$~km
significantly modifies the neutrinos' spectra,
\todo{demonstrate}
and the extended envelope outside the shock becomes a source of
higher-than-average-energy neutrinos.

The 1D matter profile and M1 transport evolution are computed using
open source supernova evolution code \lstinline{GR1D}
\footnote{\url{http://www.gr1dcode.org}}
\cite{ocon2010-gr1d, ocon2015-gr1d_with_nu},
using a progenitor profile from \cite{woos1995-sn_progenitors}.
The matter is described by the LS180 equation of state \cite{latt1991-nuc_eos},
and the opacities are computed and stored in a table
as described in Sec.~\ref{ssec:test_equilibrium}.
This standard test is also presented for example in
\cite{ocon2015-gr1d_with_nu,fouc2015-m1_nsbh,abdi2012-monte_carlo}.
\todo{also lieb2005, mull2010}

The matter profile and background field are stored on a spherical
pseudo-spectral grid composed of 11 spherical-shell subdomains
\todo{cite early SpEC paper}
comprising a total of 62 radial grid points spaced approximately
logarithmically across $r\in(0,740)$~km.
The background scattering field $\Phi(\varepsilon)$ is supplied by
\lstinline{GR1D} in the form of $J(\varepsilon_i)$,
with $\varepsilon_i$ representing 18 energy groups
identical to those in the \lstinline{NuLib} table described in
Sec.~\ref{ssec:test_equilibrium}.
For the \emph{spectral} method we use $J(\varepsilon_i)$ directly,
using zeroth-order interpolation between energy groups;
for the \emph{gray} method we use $J=\sum J(\varepsilon_i) \Delta\varepsilon_i$,
and $G=\sum J(\varepsilon_i) \Delta\varepsilon_i/\varepsilon_i$ instead,
with $\Delta\varepsilon_i$ the bin-width of the $i$th energy group.

For fiducial neutrino distributions we use the matter profile
as input into a Monte Carlo radiation transport calculation using
open source neutrino transport code \lstinline{Sedonu}
\footnote{\url{https://bitbucket.org/srichers/sedonu}}
\cite{rich2015-monte_carlo}.
To homogenize the physics modeled across these three treatments
(M1 transport to provide the background fields,
ray tracing to compute neutrino distributions, and
Monte Carlo transport for a fiducial comparison)
we turn off inelastic scattering in the M1 and Monte Carlo codes,
since we have not incorporated that physics into this ray tracing code,
and we turn off general relativistic effects in the ray tracing code,
since the Monte Carlo code assumes Minkowski spacetime.

\begin{figure}
  \includegraphics[width=\columnwidth]{fig-collapse_profile_avg_eps_vs_cosA}
  % Data:   $DIR/
  % Script: $DIR_GRAY_SC/../plot-compare_spectral_vs_gray-angular_distribution_avg_eps-production.gnu
  % w/  DIR_MC     =$DIR/ResultsSherwood-03/Spectra
  %     DIR_SPEC_SC=$DIR/ResultsBrett-A-03/Test-scat-01
  %     DIR_GRAY_SC=$DIR/ResultsBrett-B-03/Test-scat-01
  %     DIR_GRAY_AB=$DIR/ResultsBrett-B-03/Test-noscat-01
  % and DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/CollapseProfile/
  \caption{Angular distribution of $\nu_x$ average energies
    $\langle\varepsilon\rangle(\cos A)$
    in the collapse profile test described in Sec.~\ref{ssec:test_collapse}.
    The observer is at 500~km, and the shock at 150~km,
    so that the half-opening angle of the shock is $\cos A\approx0.954$.
    The four methods depicted are 1) a fiducial Monte Carlo calculation,
    ray tracing using the 2) spectral and 3) gray methods to estimate background
    fields for scattering, and 4) ray tracing neglecting scattering.
  }
  \label{fig:avg_eps_collapse}
\end{figure}

We place a stationary observer at $r=500$~km.
Fig.~\ref{fig:avg_eps_collapse} shows the distribution of average $\nu_x$
energies across incoming angles to this observer.
We show $\nu_x$ because they present the largest scattering effects:
they scatter through a thicker atmosphere outside their deep emission surface,
and their hotter spectrum experiences stronger modification due to the
$\varepsilon^2$ dependence of the scattering cross-section.
Against the fiducial Monte Carlo distribution, we show the
\emph{noscat} treatment, and the \emph{scat} treatment using both the
\emph{spectral} method and the \emph{gray} method.

As expected the average energies from both
the scattering envelope ($\cos A \lesssim 0.95$) and
the bright core ($\cos A\gtrsim 0.995 $)
are well characterized by the \emph{scat} treatments
and badly characterized by the \emph{noscat} treatment.
In this case for $\nu_x$ the major effect of elastic scattering
is to decrease the average energy of neutrinos coming from the core
and increase the average energy of neutrinos coming from the envelope.

We don't show it here, but the angular distribution contributing to the
\todo{do show it}
total number density is also strongly affected by elastic scattering.
Without elastic scattering, the central regions emitting 60\%
of the neutrinos for the species $\{\nu_e,\bar{\nu}_e,\nu_x\}$
have scales $R_\nu\approx\{40,35,20\}$~km;
with elastic scattering the scales are
$R_\nu\approx\{50,45,35\}$~km.
% From by-hand summation over histogram G(cosA_i)*dcosA
% Script: $DIR/plot-compare_scat_vs_noscat-G_angular_distribution-all.gnu
% DIR=.../20161007-ray_tracing_tests/CollapseProfile/ResultsBrett-B-03

Total luminosities and average energies from the different treatments are
\todo{update MC results 02->03}
presented in Tab.~\ref{tab:collapse_rt_vs_mc}.
As expected, $\nu_e$ is least affected by scattering, and $\nu_x$ most.
In fact, without scattering, $\nu_x$ luminosities are overestimated by more than
two orders of magnitude, due to the steep increase in temperature with depth
in the inner core.
Though contributing only a fraction of the total luminosity,
the average energy of neutrinos scattered to the observer from the envelope
outside the shock, $\langle\varepsilon\rangle^{\rm sc}$,
is poorly characterized by the \emph{noscat} treatment for all species.
By contrast, both the \emph{gray} and \emph{spectral} \emph{scat} treatments
faithfully describe the high scattered energies from the envelope.

\begin{table}%[h]
  % to get mc luminosities: tail $MCDATA/Spectra/luminosities_and_avg_eps-501km-nu?.dat
  % to get rt luminosities: tail $RTDATA/Moments-nu?.out
  % MCDATA=$DIR/ResultsSherwood-02
  % RTDATA_NO=$DIR/ResultsBrett-B-03/Test-noscat-01/
  % RTDATA_GR=$DIR/ResultsBrett-B-03/Test-scat-01/
  % RTDATA_SP=$DIR/ResultsBrett-A-03/Test-scat-01/
  % DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20161007-ray_tracing_tests/CollapseProfile/
  \caption{
    Comparison of total luminosities and average energies between
    the following methods:
    ray tracing ignoring elastic scattering `RT noscat',
    ray tracing with the gray scattering treatment `RT scat gr',
    ray tracing with the spectral scattering treatment `RT scat sp', and
    a fiducial Monte Carlo transport evolution `MC'.
    We also show average energies of the scattering envelope
    $\langle\varepsilon\rangle^{\rm sc}$, which are estimated by eye
    from plots like Fig.~\ref{fig:avg_eps_collapse} at $\cos A < 0.95$.
    Luminosities have units $10^{52}\,{\rm erg}\,{\rm s}^{-1}$ and
    average energies have units MeV.
    The $\nu_x$ luminosities are per-species:
    multiply by four to get the total heavy-lepton neutrino luminosities.
  }
  \label{tab:collapse_rt_vs_mc}
  \begin{tabularx}{\columnwidth}{X r r r r}
    & {\bf RT noscat} & \,\,{\bf RT scat gr} & \,\,{\bf RT scat sp} & \,\,{\bf MC} \\
    \hline
    $L_{\nu_e}$                                         &  3.53 & 3.14 & 3.04 & 3.48 \\
    $L_{\bar{\nu}_e}$                                   & 5.19 & 3.05 & 3.03 & 3.01 \\
    $L_{\nu_x}$                                         & 222.0  & 1.88 & 1.76 & 1.70 \\
    $\langle \varepsilon_{\nu_e} \rangle$               &  10.6 & 10.6 & 10.9 & 11.0 \\
    $\langle \varepsilon_{\bar{\nu}_e} \rangle$         & 14.2 & 13.0 & 13.8 & 13.7 \\
    $\langle \varepsilon_{\nu_x} \rangle$               & 47.8 & 16.0 & 17.3 & 16.2 \\
    $\langle \varepsilon_{\nu_e} \rangle^{\rm sc}$      & 4    & 14   & 17   & 16 \\
    $\langle \varepsilon_{\bar{\nu}_e} \rangle^{\rm sc}$& 4    & 24   & 20   & 18 \\
    $\langle \varepsilon_{\nu_x} \rangle^{\rm sc}$      & 2    & 28   & 31   & 27 \\
    \hline
  \end{tabularx}
\end{table}

We can make some quantitative sense of the scattered energies in
Fig.~\ref{fig:avg_eps_collapse} and Tab.~\ref{tab:collapse_rt_vs_mc}
using the solutions explored in the simple configurations above.
In particular, the average energy in the scattering envelope
($\cos A\lesssim 0.95$) is related to the spectrum in the direction
of the interior ($\cos A \sim 1$).
The relation may be derived by simplifying our realistic model
to that of a homogeneous matter profile and background scattering field.

By expanding the exponential of
Eqns.~\ref{eqn:fse_homogeneous_general} and \ref{eqn:fae_homogeneous_general}
as we did for Eqn.~\ref{eqn:fae_homogeneous_limits},
and furthermore factoring out the dominant energy dependence from the
cross-sections (i.e.\ $\chi\equiv\varepsilon^2 \zeta$
with $\zeta$ approximately constant), we have
\begin{align}
  \label{eqn:fae_homogeneous_thin_limit}
  f_{\rm AE}(\varepsilon)
  &= \varepsilon^2 \zeta_a \,f^{\rm eq}(\varepsilon)\,s, \\
  \label{eqn:fse_homogeneous_thin_limit}
  f_{\rm SE}(\varepsilon)
  &= \varepsilon^2 \zeta_s \,\Phi(\varepsilon)\,s,
\end{align}
where $s$ is the length of scattering envelope passed through by the ray.
\todo{should be $\chi_a^*=\varepsilon\zeta_a$ for $\nu_x$}
In the envelope the local temperature is low so $f^{\rm eq}/\Phi\ll 1$,
and also $\zeta_a/\zeta_s\ll 1$,
so $f_{\rm SE}$ strongly dominates over $f_{\rm AE}$.
The average energy of the scattered field is therefore
\begin{equation}
  \label{eqn:avg_eps_sc}
  \langle \varepsilon \rangle^{\rm sc}
  = \frac{\int d\varepsilon\,\varepsilon^3 f}{\int d\varepsilon\,\varepsilon^2 f}
  \approx \frac{\int d\varepsilon\,\varepsilon^5 \Phi}{\int d\varepsilon\,\varepsilon^4 \Phi}.
\end{equation}

The spectrum of the background field $\Phi(\varepsilon)$ is well-approximated
as the \emph{scat} solution for $\cos A\sim 1$, since that is the dominant
source direction for neutrinos.
And because the scattering envelope is optically thin at all energies,
we can assume the neutrino spectrum is essentially unchanged in its passage
through the envelope.

In order to estimate $\langle \varepsilon \rangle^{\rm sc}$ analytically,
we must write the background field $\Phi(\varepsilon)$ analytically.
Direct Fermi-Dirac fits using a temperature and chemical potential
representative of a physical neutrinosurface fair poorly since neutrinos of
different energies decouple from the matter at different radii, over which
the thermodynamic state varies substantially.
Phenomenological Fermi-Dirac fits work well;
but so do pinched spectral fits which are much simpler
\cite{keil2003-pinched_spectra,miri2016-sn_neutrinos}:
\begin{equation}
  \label{eqn:pinched_spectra}
  \Phi^{\rm pi}(\varepsilon) \propto
  \varepsilon^{\alpha-2}
  \exp\left(-(\alpha+1)\frac{\varepsilon}{\langle\varepsilon\rangle}\right),
\end{equation}
(where our definition differs from that of \cite{keil2003-pinched_spectra}
by a factor of
$\varepsilon^2$, since we define our distribution function to be dimensionless
in natural units $\{\hbar,c\}=1$).

Pinching parameters
(calculated by eye from the \emph{spectral} \emph{scat} method)
for the species
$\{\nu_e,\bar{\nu}_e,\nu_x\}$ are $\alpha\approx\{3.6,5.1,2.3\}$.
Energy moments of pinched spectra like those in Eqn.~\ref{eqn:avg_eps_sc}
have a simple analytic form so that Eqn.~\ref{eqn:avg_eps_sc} becomes
\begin{align}
  \label{eqn:pinched_avg_eps_sc_1}
  \langle\varepsilon\rangle^{\rm sc}
  &\approx \frac{\alpha+3}{\alpha+1} \langle\varepsilon\rangle, \\
  %\label{eqn:pinched_avg_eps_sc_2}
  %&= \{1.4,1.3,1.6\} \langle\varepsilon\rangle, \\
  \label{eqn:pinched_avg_eps_sc_3}
  &\approx \{16,18,28\}\,~{\rm MeV},
\end{align}
again for $\{\nu_e,\bar{\nu}_e,\nu_x\}$ respectively.
These analytic predictions agree with the average energies of the scattering
envelope to approximately 10\%
% From by-eye examinations of <eps>(cosA) plot using data from
% MCDATA=$DIR/ResultsSherwood-02
% RTDATA_GR=$DIR/ResultsBrett-B-03/Test-scat-01/
% RTDATA_SP=$DIR/ResultsBrett-A-03/Test-scat-01/
%          nue     nua     nux
of all of the treatments including elastic scattering presented
in Tab.~\ref{tab:collapse_rt_vs_mc}, except for
$\bar{\nu}_e$ in the \emph{gray} \emph{scat} treatment,
which deviates from our prediction by 30\%.
This agreement is excellent despite the drastic
simplifications used in our toy model.

The $\langle\varepsilon_{\bar{\nu}_e}\rangle^{\rm sc}$ prediction
in the \emph{gray} \emph{scat} treatment is approximately 30\%
larger than the the fiducial Monte Carlo estimate. This is
due to the large negative local chemical potential
$\eta_{\bar{\nu}_e}\sim-10$ in the scattering envelope.
As described in Sec.~\ref{ssec:sources_se},
we use $\eta_\nu$ in the \emph{gray} treatment to construct our
synthetic spectrum from total neutrino densities $J$ and $G$.
This error and our successful analysis using pinched spectra above,
points the way to future improvements to the \emph{gray} method:
making better assumptions about the spectra which don't depend upon
local fluid quantities.

\subsection{Hypermassive Neutron Star Remnant:
  Testing Everything}
\label{ssec:test_disk_comparison}
The merger of two neutron stars by gravitational wave emission produces a
postmerger configuration composed of a single neutron star surrounded by a disk.
Because of strong differential rotation, the remnant temporarily avoids
collapse to a black hole, even if its mass exceeds the threshold of dynamical
instability for a rigidly rotating neutron star \cite{duez2009-review}.
These objects, called hypermassive neutron stars,
may avoid collapse for $X\textrm{--}Y$~s,
\todo{cite $X$, $Y$}
depending upon a number of physical factors including thermal pressure,
magnetic fields, and the microphysics of the fluid.

Such a configuration was modeled in \cite{fouc2016-m1_nsns} by
evolving fluid and spacetime through the
final inspiral and merger of two identical neutron stars of
isolated gravitational mass 1.2~$M_{\odot}$. The postmerger configuration
was simulated using a gray M1 transport scheme for the neutrinos, evolving the
energy density, number density, and energy flux \cite{fouc2016-m1_evolve_n}
for $\sim10\,{\rm ms}$ following merger.
The fluid was modeled using the LS220 equation of state \cite{latt1991-nuc_eos}.

We use a single time snapshot from that configuration
at $t=11\,{\rm ms}$ after merger.
The volume data are represented on a pseudospectral domain of a
sphere with concentric shells extending to $r\approx1400$~km.
Radial grid spacings are $\Delta r \approx 0.15$~km in the star
and $\Delta r \approx 2.5$~km in the disk,
with 12 cells spanning polar angles $\theta\in[0,\pi]$
and 24 cells spanning azimuthal angles $\phi\in[0,2\pi]$.
\todo{way too-low angular resolution}
Though a pseudospectral representation of non-smooth hydrodynamic data
introduces some Gibbs-like oscillations in the variables,
we choose to use this representation instead of the mesh-refined
finite difference grid of the evolution because the pseudospectral
representation uses much less memory:
95~MB in pseuspectral vs.\ 3~GB in finite difference representation.
Obviously, memory loads of this order are not insurmountable;
but they would require some modifications to our
volume data interpolation infrastructure.
For the purposes of this analysis the pseudospectral representation is adequate.
Figs.~\ref{fig:nsns_rho_merid}--\ref{fig:nsns_temp_equat}
show slices of density and temperature from the finite difference volume data.

\begin{figure}
  \includegraphics[width=\columnwidth]{production-colormap-merid-rho}
  % VTK Data: zosima:/home/mbdeaton/Research/20170602-nsns_vol_data
  % Plots:    $DIR/production-colormap-{merid,equat}-{rho,temp}.png
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170420-nsns_volume_data/02-two_domains_avg_eps/PlotsEtc
  \caption{A meridional slice of density in the hypermassive neutron
    star and disk configuration.
    The distorted rectangular boundaries are the boundaries of the grid
    used in the numerical simulation, which employs a coordinate mapping to
    concentrate points near the central object.}
  \label{fig:nsns_rho_merid}
\end{figure}

\begin{figure}
  \includegraphics[width=\columnwidth]{production-colormap-equat-rho}
  % see notes from production-colormap-merid-rho
  \caption{An equatorial slice of density in the hypermassive neutron
    star and disk configuration.}
  \label{fig:nsns_rho_equat}
\end{figure}

\begin{figure}
  \includegraphics[width=\columnwidth]{production-colormap-merid-temp}
  % see notes from production-colormap-merid-rho
  \caption{A meridional slice of temperature in the hypermassive neutron
    star and disk configuration.}
  \label{fig:nsns_temp_merid}
\end{figure}

\begin{figure}
  \includegraphics[width=\columnwidth]{production-colormap-equat-temp}
  % see notes from production-colormap-merid-rho
  \caption{An equatorial slice of temperature in the hypermassive neutron
    star and disk configuration.}
  \label{fig:nsns_temp_equat}
\end{figure}

Opacities are computed from the LS220 equation of state using
\lstinline{NuLib} and are stored as a table covering
energy, density, temperature, and composition ranges
spanning $\varepsilon\in[1,280.5]$~MeV logarithmically,
$\varrho\in[10^6,3.2\times10^{15}]\,{\rm g}\,{\rm cm}^{-3}$ logarithmically,
$T\in[0.05,150]$~MeV logarithmically, and 
$Y_e\in[0.035,0.55]$ linearly,
with grid extents$\{18,82,65,51\}$ respectively.

We place stationary (i.e.\ Eulerian) observers at fixed coordinate radius
$r=250$~km in the $y$-$z$ plane along an arc at $N_\theta$ positions
distributed linearly in $\cos\theta$ over the northern hemisphere
using $\theta_i=\cos^{-1}\left(1-i/(N_\theta-1)\right)$
for $i=0,1,2,\ldots N_\theta-1$.
We choose $N_\theta=30$.

Though the ray tracing sampling of $f$ is done in full general relativity,
in our calculation of moments at the observer positions,
we make the simplifying assumption of Minkowski spacetime.
This assumption introduces errors in our moments of order
$GM/r=4\,{\rm km}/250\,{\rm km}$
\todo{too simplistic}
(with $G$ the gravitational constant and $M$ the central mass),
or $\sim 1\%$.

Each observer samples the distribution function over a uniform grid
in energy, cosine of polar angle, and azimuthal angle with extents
$\{N_{\tilde{\varepsilon}},N_A, N_B\}$ spanning the ranges
$\tilde{\varepsilon}\in(0,100)$~MeV,
$\cos A\in\left((\cos A)_{\rm min}, 1\right)$, and
$B\in[0,2\pi)$.
The integrals for fluxes of number and energy
(Eqns.~\ref{eqn:Ha}, \ref{eqn:Ka}, and \ref{eqn:J_H_S_eps_integrated})
become sums over all rays.
Taking these fluxes in the radial direction,
and using the midpoint rule to convert the integral to a sum, we have
\begin{align}
  \label{eqn:Kr_minkowski}
  \tilde{K}^r &= \Delta
  \sum\limits_{m=0}^{N-1}
  \tilde{\varepsilon}_m^2 \cos A_m f_m,\\
  \label{eqn:Hr_minkowski}
  \tilde{H}^r &= \Delta
  \sum\limits_{m=0}^{N-1}
  \tilde{\varepsilon}_m^3 \cos A_m f_m,
\end{align}
with $f_m \equiv f(\tilde{\varepsilon}_m,\cos A_m, B_m)$,
$\Delta \equiv \Delta\tilde{\varepsilon}\, \Delta(\cos A)\, \Delta B$,
$N \equiv N_{\tilde{\varepsilon}}\, N_A\, N_B$,
and $m$ the index labeling each ray.
Average energies are given by
$\langle\tilde{\varepsilon}\rangle(\theta)=H^r(\theta)/K^r(\theta)$.
We choose $N_{\tilde{\varepsilon}}=30$, $N_A=120$, and $N_B=20$.
For the \emph{scat} case we choose $(\cos A)_{\rm min}=0.81$;
For the \emph{noscat} case, presenting a smaller emitting region,
we choose $(\cos A)_{\rm min}=0.91$.

We also combine measurements from all observers to estimate
total luminosities and averages over the sky.
We assume the disk is azimuthally symmetric and
symmetric with respect to reflection across the equatorial plane.
Figs.~\ref{fig:nsns_rho_merid} and \ref{fig:nsns_rho_equat} show how these
assumptions are approximately valid at 11~ms after merger.
Total luminosities are then computed as integrals of the radial fluxes
over $\cos\theta$.
Using the trapezoid rule the number and energy luminosities become sums:
\begin{align}
  \label{eqn:luminosity_R}
  R &= C
  \left(\frac{1}{2}(\tilde{K}^r_0+\tilde{K}^r_{N_\theta-1})
  +\sum\limits_{m=1}^{N_\theta-2}\tilde{K}^r_m\right),\\
  \label{eqn:luminosity_L}
  L &= C
  \left(\frac{1}{2}(\tilde{H}^r_0+\tilde{H}^r_{N_\theta-1})
  +\sum\limits_{m=1}^{N_\theta-2}\tilde{H}^r_m\right),
\end{align}
with $C \equiv 4\pi r^2\, \Delta(\cos \theta)$,
\todo{source of GR error in coordinate $r$}
$\tilde{K}^r_m \equiv \tilde{K}^r(\theta_m)$, and
$\tilde{H}^r_m \equiv \tilde{H}^r(\theta_m)$.
Average energies over the whole sky are then
$\langle\tilde{\varepsilon}\rangle=L/R$.

\begin{table}%[h]
  % Total fluxes over the sky integrated in google sheet: 20170726-theta_distribution_integrations
  % RTDATA_NO=$DIR/PositionDistributions-nu?-250km-Lev1-e30a120b20-noscat-01.dat
  % RTDATA_GR=$DIR/PositionDistributions-nu?-250km-Lev1-e30a60b10-scat-01.dat
  % DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170324-ray_tracing_nsns/Run03/PositionDistributions
  \caption{
    Comparison of total luminosities and average energies
    of the hypermassive neutron star configuration.
    The methods are ray tracing using the \emph{noscat} method `RT noscat',
    ray tracing using the \emph{gray} \emph{scat} method `RT scat',
    and the M1 transport simulation `M1'.
    The ray tracing totals were computed from sums over
    observers placed in the $y$-$z$ plane.
    M1 values are taken from \cite[Figs. 7, 9, 10]{fouc2016-m1_evolve_n}.
    Energy luminosities have units $10^{52}\,{\rm erg}\,{\rm s}^{-1}$,
    number luminosities $10^{57}\,{\rm s}^{-1}$, and
    average energies MeV.
    The $\nu_x$ luminosities are per-species:
    multiply by four to get the total heavy-lepton neutrino luminosities.
  }
  \label{tab:nsns_rt_vs_m1}
  \begin{tabularx}{\columnwidth}{X r r r}
    & {\bf RT noscat} & \,\,{\bf RT scat} & \,\,{\bf M1} \\
    \hline
    $L_{\nu_e}$                                 & 5.62 & 6.05 & 5 \\
    $L_{\bar{\nu}_e}$                           & 10.6 & 11.8 & 12 \\
    $L_{\nu_x}$                                 & 23.3 & 12.7 & 12 \\
    $R_{\nu_e}-R_{\bar{\nu}_e}$                 & -1.22 & -1.85 & -2 \\
    $\langle \varepsilon_{\nu_e} \rangle$       & 12.08 & 11.91 & 12 \\
    $\langle \varepsilon_{\bar{\nu}_e} \rangle$ & 16.07 & 14.67 & 15 \\
    $\langle \varepsilon_{\nu_x} \rangle$       & 34.09 & 23.58 & 26 \\
    \hline
  \end{tabularx}
\end{table}

Tab.~\ref{tab:nsns_rt_vs_m1} compares the all-sky
\todo{get accurate M1 numbers from Francois}
luminosities and average energies from ray tracing and from
the M1 transport simulation,
which serves as our fiducial calculation.
Uncertainties are around 15\% for energy luminosities and
10\% for average energies from the M1 simulation
\cite[Sec.~A.6]{fouc2016-m1_evolve_n}.
Because the model was evolved with a gray M1 transport scheme,
the only \emph{scat} method we can use in our ray tracing is the
\emph{gray} method.
As in the post-bounce configuration in Sec.~\ref{ssec:test_collapse},
the \emph{scat} treatment is more faithful than the \emph{noscat} treatment,
significantly so for the heavy-lepton neutrinos.
(Though, we note the energy luminosity for $\nu_e$ is overestimated
20\% by both \emph{noscat} and \emph{scat} treatments,
and the overestimate is slightly worse for the \emph{scat} treatment.)
\todo{explain}
With scattering turned off ray tracing overestimates $L_{\nu_x}$ by 200\%;
but with scattering turned on agreement in all measures
is better than 10\%, except for $L_{\nu_e}$ which is overestimated by 20\%.
As the M1 transport evolution also demonstrates,
$R_{\bar{\nu}_e}$ dominates over $R_{\nu_e}$
\todo{Need $R$ magnitude not just difference}
because the disrupted neutron star material is releptonizing.

\begin{figure}
  \includegraphics[width=\columnwidth]{theta_distrib-250km-Kr}
  % Data_a:   $DIR/PositionDistributions/*Lev1-e30a120b20-01*
  % Data_s:   $DIR/PositionDistributions/*Lev1-e30a60b10-01*
  % Data gathered using: $DIR/CollectPositionDistributions.sh
  % Script: $DIR/GnuScripts/plot-theta_distributions-Hr-fixed_r.gnu
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170324-ray_tracing_nsns/Run03
  \caption{Radial number fluxes, $K^r(\theta)$,
    as defined in Eqn.~\ref{eqn:Kr_minkowski},
    in the hypermassive neutron star configuration.
    Sampled for observers at fixed coordinate distances from the center
    of the star, with $r=250\,{\rm km}$.
  }
  \label{fig:nsns_theta_distrib_Kr}
\end{figure}

\begin{figure}
  \includegraphics[width=\columnwidth]{theta_distrib-250km-Hr}
  % see notes from theta_distrib-250km-Kr
  % Script: $DIR/GnuScripts/plot-theta_distributions-Hr-fixed_r.gnu
  \caption{Same as Fig.~\ref{fig:nsns_theta_distrib_Kr} but radial energy
    fluxes $H^r(\theta)$ as defined in Eqn.~\ref{eqn:Hr_minkowski}.
  }
  \label{fig:nsns_theta_distrib_Hr}
\end{figure}

The distributions of radial fluxes over observer position are shown
in Figs.~\ref{fig:nsns_theta_distrib_Kr} and \ref{fig:nsns_theta_distrib_Hr}.
With the \emph{noscat} treatment the $\nu_x$ luminosities are relatively
constant in $\theta$, because the disk is optically thin to $\nu_x$,
and most of the neutrinos come from the hypermassive neutron star,
which is roughly spherical.
The slight upward trend in $\nu_x$ luminosities with $\theta$ may be due to
asymmetries in the configuration.
It is not due to the Doppler shift from the rapid rotation of the star
(a hypothesis we tested by setting $u_i=0$ and $W=1$ in the ray
tracing equations).
Because of their larger absorption cross section,
the disk is not optically thin to $\nu_e$ and $\bar{\nu}_e$,
and observers within view of the hot hypermassive neutron star measure
the largest radial fluxes.

When we turn on scattering, the disk is no longer transparent to $\nu_x$,
and $\nu_x$ luminosities present the same qualitative $\theta$-dependence
as that of the other species.
In general, scattering increases luminosities for polar observers
($\theta \sim 0^\circ$) because additional high energy neutrinos
from the hypermassive neutron star are scattered by the disk back up to
the observer.
For observers in the equatorial plane ($\theta \sim 90^\circ$)
the effect of scattering is a decrease in energy luminosities,
since some of the highest energy neutrinos from deep in the interior
are scattered away from the line of sight by the disk.

\begin{figure}
  \includegraphics[width=\columnwidth]{theta_distrib-250km-avg_eps}
  % see notes from theta_distrib-250km-Kr
  % Script: $DIR/GnuScripts/plot-theta_distributions-avg_eps_from_fluxes-fixed_r.gnu
  \caption{Same as Fig.~\ref{fig:nsns_theta_distrib_Kr} but
    average energies of radial fluxes
    $\langle\tilde{\varepsilon}\rangle(\theta)=H^r(\theta)/K^r(\theta)$,
    as defined in Eqns.~\ref{eqn:Kr_minkowski} and \ref{eqn:Hr_minkowski}.
  }
  \label{fig:nsns_theta_distrib_avg_eps}
\end{figure}

Fig.~\ref{fig:nsns_theta_distrib_avg_eps} shows the distribution of
average energies of radial fluxes over observer positions.
With and without scattering, the average energies are highest
for observers near the polar axis, since polar observers get a
direct view of the hot hypermassive neutron star.
Scattering decreases average energies across all observer positions,
as it did in the post-bounce configuration, or any configuration of
a hot interior surrounded by a scattering envelope.
The strength of the effect of scattering on the average energies of
the different species is seen to follow the ranking
$\nu_x>\bar{\nu}_e>\nu_e$ due to two factors:
the average energies of the spectra obey the same ranking,
and the thicknesses of the different species' scattering envelopes
also obey the same ranking.
As we showed in the optically thin limit of
Eqn.~\ref{eqn:fse_homogeneous_thin_limit},
\todo{too simplistic}
the scattering contribution $f_{\rm SE}$ is
proportional to energy $\varepsilon^2$ and path length $s$.

In Figs.~\ref{fig:nsns_cosA_distrib_I_00deg}
and \ref{fig:nsns_cosA_distrib_I_90deg}, we show the distributions of neutrino
number density over incoming angle
for the observer on the polar axis $\theta=0^\circ$,
and in the equatorial plane $\theta=90^\circ$,
as defined in Eqn.~\ref{eqn:J_per_cosA}.
As in the case of the collapse profile in Sec.~\ref{ssec:test_collapse},
the dominant effect of \emph{scat} on $\tilde{G}(\cos A)$
is to spread the distribution out to larger angles.
\todo{also examine spectra or average energies}

\begin{figure}
  \includegraphics[width=\columnwidth]{cosA_distrib-intensity-250km-00deg}
  % Data:   $DIR/Lev1-e30a60b10-{,no}scat-00
  % Script: $DIR/plot-angular_distribution-intensity-compare.gnu
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170324-ray_tracing_nsns/Run03/000Mo-000.0deg-250km
  \caption{Angular distributions of fluxes for an observer on the rotation axis
    in the hypermassive neutron star configuration.}
  \label{fig:nsns_cosA_distrib_I_00deg}
\end{figure}

\begin{figure}
  \includegraphics[width=\columnwidth]{cosA_distrib-intensity-250km-90deg}
  % Data:   $DIR/Lev1-e30a60b10-{,no}scat-00
  % Script: $DIR/plot-angular_distribution-intensity-compare.gnu
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170324-ray_tracing_nsns/Run03/000Mo-090.0deg-250km
  \caption{Angular distributions of fluxes for an observer in the equatorial
    plane in the hypermassive neutron star configuration.}
  \label{fig:nsns_cosA_distrib_I_90deg}
\end{figure}

\section{Applications}
\label{sec:applications}

\subsection{Neutrino Oscillations at High Neutrino Densities}
\label{ssec:V_nunu}
The Boltzmann Equation (Eqn.~\ref{eqn:boltzmann}) is one limiting
form of the quantum kinetic equations \cite{vlas2014-neutrino_qkes},
the limit where collisional mean free paths are much shorter
than oscillation lengths (i.e.\ $\varepsilon/\mathscr{K} \ll |H|^{-1}$
with $H$ the Hamiltonian describing coherent forward scattering interactions).
In this limit the neutrino density matrix takes the form
\begin{equation}
  \label{eqn:density_rho_0}
  \rho_0(\varepsilon,\ell_\mu) =
  \sum\limits_{\alpha}f^{\nu_\alpha}(\varepsilon,\ell_\mu)
  \ket{\nu_\alpha} \bra{\nu_\alpha},
\end{equation}
with $\ket{\nu_\alpha}$ the neutrino flavor eigenstates.
The neutrino state remains pure, or diagonal in the flavor basis.

But at the opposite limit, in the free-streaming regime,
coherence effects become important,
and the quantum kinetic equations take a Schr\"{o}dinger-like form:
\begin{equation}
  \label{eqn:schroedinger}
  i \frac{d}{ds} S = H S,
\end{equation}
where $ds$ is an interval of proper length, and
\todo{mention $\bar{\nu}$ eqn}
$S$ is the flavor evolution matrix describing a mixed state
\begin{equation}
  \label{eq:density_evolution}
  \rho = S \, \rho_0 \, S^\dagger.
\end{equation}
In this limit, we may decompose the Hamiltonian into vacuum,
matter, and neutrino contributions:
\begin{equation}
  \label{eqn:decompose_H}
  H = H_{\rm V} + H_e + H_{\nu\nu}.
\end{equation}
Explicit formulas for these terms may be found in the oscillation literature,
e.g.\ \cite{duan2009-review},
but for clarity we give their order-of-magnitude scales here:
\begin{align}
  \label{eqn:scale_H_V}
  |H_{\rm V}| &\sim \frac{\Delta m^2}{\varepsilon}, \\
  \label{eqn:scale_H_e}
  |H_e|       &\sim G_{\rm F}\left|n_{e^-}-n_{e^+}\right|, \\
  \label{eqn:scale_H_nunu}
  |H_{\nu\nu}|&\sim G_{\rm F} \left|G_{\nu_e}-G_{\bar{\nu}_e}\right|,
\end{align}
where $\Delta m$ is the mass difference between neutrino mass eigenstates,
$G_{\rm F}$ is the Fermi coupling constant,
$n_{e^-}$ and $n_{e^+}$ are the electron and positron number densities,
and $G_{\nu_e}$ and $G_{\bar{\nu}_e}$ the electron neutrino and antineutrino
number densities defined in Eqn.~\ref{eqn:G}.
For the scale of $H_{\nu\nu}$ we have only included the isotropic
components of the neutrino fields.

In regimes in which $|H_{\nu\nu}|\ll|H_{\rm V},H_e|$,
the flavor evolution along the ray is soluble, and reveals a rich
and physically important phenonenology including
vacuum, solar, atmospheric, and terrestrial oscillations,
as well as oscillations in supernova envelopes.
Where neutrino densities are relatively high, however,
as in neutron star mergers, the problem must be solved globally.
\todo{be clearer}
To date, no method has been devised to handle this problem.

However, we can solve a similar but tractable problem along a single ray.
Assume all the neutrino rays intersecting an event along a given test ray
have undergone the same flavor evolution history as that of the test ray,
i.e.\ $S$ is the same for all rays sharing that event.
This is the so-called single-angle approximation.
\todo{cite someone}
\todo{say something about its validity}
Though widely used in the supernova oscillation literature
and shown to be qualitatively faithful in those environments,
\todo{cite vlasenko?}
a recent study has shown that the single-angle approximation
may mask certain instabilities in the flavor evolution
\cite{wu2017-fast_neutrino_conversions}.

Within the single-angle approximation,
solving for the flavor evolution history along a given test ray
requires knowledge of the unoscillated neutrino contribution to the Hamiltonian
along that trajectory.
This is a function of the unoscillated neutrino self-interaction potential,
which for the $\alpha$-th flavor is:
\begin{equation}
  \label{eqn:V_nu_alpha}
  V_{\nu_\alpha,0}(\varepsilon,\ell_\mu) =
  \frac{\sqrt{2}G_{\rm F}}{(2\pi)^3} \varepsilon^2
  \oint d\Omega'(1-\varpi')f^{\nu_\alpha}(\varepsilon,\ell'_\gamma),
\end{equation}
with the test ray propagating in direction $\ell_\mu$,
ambient rays propagating in directions $\ell'_\gamma$,
and the cosine of the angle between these given by
$\varpi'=\psi^{\mu\gamma} \ell_\mu \ell'_\gamma$.
As in the moment equations (Eqns.~\ref{eqn:J}-\ref{eqn:Ka}),
the integral is taken over all directions $\ell'_\gamma$.

If conditions are right, a resonant flavor transition
introducing significant mixing may occur very near
the point where neutrinos begin free-streaming.
The matter-neutrino resonance
\cite{malk2012-mnr_1,malk2015-mnr_2,malk2016-mnr_3},
can occur where the matter potential and
neutrino self-interaction potentials cancel,
or where $V_e - V_{\nu\nu,0}=0$, with
\begin{align}
  \label{eqn:V_e}
  V_e &= \sqrt{2} G_{\rm F} \frac{\varrho}{m_N} Y_e, \\
  \label{eqn:V_nunu0}
  V_{\nu\nu} &= \int d\varepsilon
  \big(V_{\nu_e,0}(\varepsilon)-V_{\bar{\nu}_e,0}(\varepsilon)\big),
\end{align}
with $\varrho$ the rest density and $m_{N}$ the nucleon mass.
The matter potential is always positive;
and outside the accretion disks formed in neutron star mergers,
in which the disrupted neutron-star matter is rapidly releptonizing,
the neutrino self-interaction potential is large and negative.

We examine this effect in the post-merger configuration already analyzed
in Sec.~\ref{ssec:test_disk_comparison}.
We calculate the self-interaction potential at $N_s=5$ points between
$r\in[30,135]$~km,
by sampling distribution functions over a grid with extents
$N_\varepsilon=30$, $N_A=30$, $N_B=10$,
over all angles and an energy range of $\varepsilon\in(0,100)$~MeV.
\todo{improve resolution and update}
We interpolate the logarithm of the self-interaction potentials
(Eqn.~\ref{eqn:V_nu_alpha})
in path length $\log s$ along the ray,
by fitting a 3rd order spline with continuous derivative,
across the $N_s$ observation points.
The matter potential along this test ray is from an analytic wind
model qualitatively consistent with the densities in the
simulated volume (i.e.\ inside $r\sim70$~km).

In Figs.~\ref{fig:V_nunu-noscat} and \ref{fig:V_nunu-scat}
we show the self-interaction potential and its contributions from
$\nu_e$ and $\bar{\nu}_e$
for a test neutrino ray moving out along a coordinate radial trajectory
from the hypermassive neutron star used in Sec.~\ref{ssec:test_disk_comparison},
along an angle of $\theta=22^\circ$ with respect to the polar axis.
\todo{note coordinate ray}
Fig.~\ref{fig:V_nunu-noscat} shows these terms for the \emph{noscat} treatment,
and Fig.~\ref{fig:V_nunu-scat} for the \emph{scat} treatment.
Obviously including the effects of elastic scattering
tends to increase the $\bar{\nu}_e$ contribution relative to the $\nu_e$
contribution, causing the self-interaction potential to stay negative
along the entire trajectory.

\begin{figure}
  \includegraphics[width=\columnwidth]{20170619-potentials-Vnue_Vnua_Vnu-000Mo-022deg-noscat}
  % Data:   $DIR/000Mo-022deg-*km/Lev1-e30a60b10-noscat-00
  % Script: $DIR/nu_potentials.nb
  % Config: $DIR/RunAB/RunAB-noscat/nu_potentials-config-000Mo-02deg.nb
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170412-self_interaction_potentials_M12_M12-LS220-M1/Run03/
  \caption{Neutrino self-interaction potential along a test trajectory
    originating at the surface of the hypermassive neutron star, and
    proceeding outward along a radial coordinate ray with angle
    $\psi=22^{\circ}$ with respect to the polar axis.
    The trajectory is parameterized by the coordinate length $s$.
    This calculation was done with scattering turned off.
    Outside of the symmetric point at $s\sim40\,{\rm km}$, the potential is
    positive, i.e.\ $\nu_e$-dominated.
    Compare to Fig.~\ref{fig:V_nunu-scat}.
    }
  \label{fig:V_nunu-noscat}
\end{figure}

\begin{figure}
  \includegraphics[width=\columnwidth]{20170619-potentials-Vnue_Vnua_Vnu-000Mo-022deg-scat}
  % Data:   $DIR/000Mo-022deg-*km/Lev1-e30a60b10-scat-00
  % Script: $DIR/nu_potentials.nb
  % Config: $DIR/RunAB/RunAB-scat/nu_potentials-config-000Mo-02deg.nb
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170412-self_interaction_potentials_M12_M12-LS220-M1/Run03/
  \caption{Same as Fig.~\ref{fig:V_nunu-noscat},
    but with scattering turned on.
    Unlike the case with scattering turned off, here the potential is
    everywhere negative, i.e.\ $\bar{\nu}_e$-dominated.
    }
  \label{fig:V_nunu-scat}
\end{figure}

We also solve for the flavor evolution of this system,
integrating $S$ according to Eqn.~\ref{eqn:schroedinger},
as described in \cite{zhu2016-mnr_nsns_remnant}.
In Figs.~\ref{fig:survival-noscat} and \ref{fig:survival-scat}
we show the survival probabilities for $\nu_e$ and $\bar{\nu}_e$
for the two treatments.
The survival probabilities are the terms of
the flavor evolution matrix $S$ that describe how much of the original flux
of a given flavor survives a transition.
\todo{define explicitly}
Where $P_{\nu_e\rightarrow\nu_e}$ vanishes, for example in
Fig.~\ref{fig:survival-noscat}, all the original $\nu_e$ content of the beam
is oscillated into $\nu_\mu$ and $\nu_\tau$; but the $\nu_e$ density doesn't
vanish altogether,
since some $\nu_\mu$ and $\nu_\tau$ oscillates into $\nu_e$.

We see that the matter-neutrino resonance occurs in both treatments,
but with the very different behaviors of symmetric and asymmetric
resonances respectively for the \emph{noscat} and \emph{scat} cases,
as seen in Figs.~\ref{fig:survival-noscat} and \ref{fig:survival-scat}.
These two classes of matter-neutrino resonance
are described for example in \cite{malk2016-mnr_3,vaan2016-uncovering_mnr}.
This dramatic difference in survival probabilities between the \emph{noscat}
and \emph{scat} cases along this test ray emphasizes the importance of
scattering physics in the problem of neutrino oscillations in
neutron star mergers.

\begin{figure}
  \includegraphics[width=\columnwidth]{fig-survival_prob-noscat_nh}
  % Flavor evolution by Yonglin, see emails 7.19.17-7.20.17
  % Plot: $DIR/31-900kmNHnoScatterPot.eps
  % w/ DIR=/Users/mbdeaton/Documents/ResearchGraveyard/NuclearAstrophysics/DataAndVis/20170412-self_interaction_potentials_M12_M12-LS220-M1/Run03/RunAB/probabilityPotential
  \caption{Neutrino survival probabilties for the \emph{noscat} case.
    In the lower panel we reproduce the unoscillated self-interaction potential
    $V_{\nu\nu,0}$ from Fig.~\ref{fig:V_nunu-noscat} and a matter potential
    computed assuming a very simple steady state wind consistent with the
    densities within the simulation volume inside of $r<70$~km
    along this trajectory.
    In this case the neutrinos undergo a symmetric MNR transition
    inside of $r<50$~km,
    with the $e$-flavor neutrinos and antineutrinos fully converting to $\nu_x$.
  }
  \label{fig:survival-noscat}
\end{figure}

\begin{figure}
  \includegraphics[width=\columnwidth]{fig-survival_prob-scat_nh}
  % see notes from fig-survival_prob-noscat_nh
  % Plot: $DIR/130-800kmNHScatterPot.eps (normal hierarchy)
  \caption{Same as Fig.~\ref{fig:survival-noscat},
    but with scattering turned on.
    In this case the neutrinos undergo a standard MNR transition
    outside of $r>200$~km,
    with the $e$-flavor neutrinos and antineutrinos
    partially converting to $\nu_x$.
    }
  \label{fig:survival-scat}
\end{figure}

%\subsection{Neutrino Annihilation Integral}
%\label{ssec:q_nunu}
%In this section we calculate $q_{\nu\bar{\nu}}$ at a representative point
%in the funnel of the NSNS configuration and show that including scattering makes
%a difference (or not) in our estimate of the power available to drive a jet.

% ******************************************************************************
\appendix

\section{Momentum Definitions}
\label{sec:def_momentum}
Here we describe our conventions defining the four-momentum of a neutrino.
We decompose the neutrino momentum into components parallel and orthogonal to
an observer's velocity $u_\beta$:
\begin{equation}
  \label{eqn:def_momentum_2}
  p_\beta = \varepsilon (u_\beta + \ell_\beta),
\end{equation}
with $u_\alpha\ell^\alpha=0$ and $\ell_\alpha\ell^\alpha=1$.

We use two possible fiducial observers to define the momentum
decomposition via Eqn.~\ref{eqn:def_momentum_2}: the Eulerian,
or normal observer $n_\mu=-\alpha \,\partial_\mu t$,
and the fluid, or comoving observer
$u^\mu=Wn^\mu+v_{\rm E}^\mu$.
Here $t$ is coordinate time and $\alpha$ is lapse
in the standard 3+1 decomposition of the metric
\begin{equation}
  \label{eqn:adm_metric}
  \psi_{\mu\gamma} \rightarrow
  \left(
  \begin{matrix}
    -\alpha^2 + \beta^i \beta_i  & \beta_i \\
    \beta_j                      & g_{ij}
  \end{matrix}
  \right).
\end{equation}
We have also introduced
the fluid Lorentz factor $W=-n_\mu u^\mu=\alpha u^t$,
its Eulerian velocity $v_{\rm E}^\mu=g^\mu_\lambda u^\lambda$
(distinct from its coordinate velocity $v^\mu=u^\mu/u^t$),
and the projection tensor orthogonal to the normal observer
$g^\mu_\lambda=\psi^\mu_\lambda+n^\mu n_\lambda$,
whose spatial part is the 3-metric, $g_{ij}$.

We specify the neutrino direction in the observer's frame
with two spherical polar angles
($a$,$b$) with respect to the simulation cartesian coordinates
\begin{align}
  \ell_\alpha &\rightarrow q(s,\Omega_i), \\
  \Omega_i &\rightarrow (\sin a \cos b,\sin a\sin b,\cos a),
\end{align}
or alternatively the two spherical polar angles
($A$,$B$) with respect to rotated coordinates
\begin{equation}
  \label{eqn:def_direction_primed_2}
  \Omega_{i'} \rightarrow  (\sin A \cos B,\sin A\sin B,\cos A).
\end{equation}
The two coordinate systems are related by a standard Euler rotation of first
$\phi$ about the $z$-axis, then $\theta$ about the rotated $y$-axis,
with $\phi$ and $\theta$ the azimuthal and polar position of the observer.
Expressed algebraically:
\begin{align}
  \label{eqn:rotation_connection}
  \Omega_x &= \Omega_{x'}\cos\theta\cos\phi-\Omega_{y'}\sin\phi+\Omega_{z'}\sin\theta\cos\phi,\nonumber\\
  \Omega_y &= \Omega_{x'}\cos\theta\sin\phi+\Omega_{y'}\cos\phi+\Omega_{z'}\sin\theta\sin\phi,\nonumber\\
  \Omega_z &= \Omega_{x'}\sin\theta+\Omega_{z'}\cos\theta.
\end{align}

The scale factors $q$ and $s$ are functions of the neutrino
direction $\Omega_i$, the observer's velocity $u_\alpha$,
and the spacetime metric.
In the case of a fluid observer, specified by an arbitrary $W$ and $u_i$,
$q$ and $s$ are given by
\begin{align}
  \label{eqn:q_general}
  q
  &= W\alpha\Bigg(2\beta^i\Omega_i W^2(\beta^i\Omega_i-1) \nonumber\\
  &\qquad\qquad - 2\Omega_i u_j g^{ij} W \alpha(\beta^i\Omega_i-1)\nonumber\\
  &\qquad\qquad + \alpha^2\left((\Omega_i u_j g^{ij})^2+\Omega_i\Omega_j g^{ij}W^2\right)\Bigg)^{-1/2}\\
  \label{eqn:s_general}
  s
  &= \beta^i\Omega_i -\Omega_i u_j g^{ij}.
\end{align}
In the case of an Eulerian observer, $W=1$ and $u_i=0$,
and these expressions simplify considerably:
\begin{align}
  \label{eqn:q_eulerian}
  \tilde{q}
  &= \alpha\left(2\beta^i\Omega_i(\beta^i\Omega_i-1)+\alpha^2\Omega_i\Omega_j g^{ij}\right)^{-1/2}\\
  \label{eqn:s_eulerian}
  \tilde{s}
  &= \beta^i\Omega_i.
\end{align}
In the even simpler case of Minkowski spacetime, these expressions
reduce to $q=W/(1+s)$, $s=-\Omega_i u^i$,
$\tilde{q}=1$, $\tilde{s}=0$.

\section{Source Terms}
\label{sec:source_terms}
Here we present the sources comprising the right hand side of the
Boltzmann Equation (Eqn.~\ref{eqn:boltzmann}). The sources for the neutrino
distribution function $f(x^\alpha;p_\beta)$ arise from collision processes
producing, removing, or scattering to/from that point in phase space.
The weak interaction rates for each process involve integrals of
the neutrino distribution function $f(x^\alpha;p'_\beta)$ and that of the
antineutrino $\bar{f}(x^\alpha;p'_\beta)$ over a momentum volume
$dP'$ (Eqn.~\ref{eqn:dP}).

We follow
\cite{brue1985-core_collapse} and \cite[Sec.~4]{shib2011-truncated_moment}
by separating these processes into four categories:
\begin{equation}
  \label{eqn:four_sources}
  C[f] \equiv C_{\rm AE} + C_{\rm SE} + C_{\rm SI} + C_{\rm PP},
\end{equation}
representing charged-current absorption and emission, elastic scattering,
inelastic scattering, and the thermal pair processes of annihilation and
production. In this work, however, we only treat absorption/emission
and elastic scattering.
We seek to write each collision source linear in $f$:
\begin{align}
  \label{eqn:sources_cae}
  C_{\rm AE}
  &= \mathscr{E}_{\rm AE} + \mathscr{K}_{\rm AE}\, f, \\
  \label{eqn:sources_cse}
  C_{\rm SE}
  &= \mathscr{E}_{\rm SE} + \mathscr{K}_{\rm SE}\, f.
\end{align}

Each term
is computed by summing the weak interaction rates of the processes from the
relevant category given in Tab.~\ref{tab:neutrino_processes}.
We compute these rates in the rest frame of the fluid,
but because they are spacetime invariants
they take the same numerical value in any frame of reference
and are completely independent of our choice of fiducial observer $u^\alpha$
(see discussion around Eqn.~\ref{eqn:boltzmann_linear}).

Each source term is unique to the neutrino or antineutrino species modeled,
and consists of a sum over all of the processes contributing to that
category of interaction. For example:
$C^{\nu_e}_{{\rm AE}} = \sum_i\,C^{\nu_e}_{{\rm AE},i}$,
where $i$ labels the absorption/emission processes involving $\nu_e$
in Tab.~\ref{tab:neutrino_processes}.
By contrast $C^{\nu_x}_{{\rm AE}} = 0$,
\todo{but we use a nonzero $C^{\nu_x}_{{\rm AE}}$}
where $\nu_x$ stands for any of the four heavy-lepton neutrinos and
antineutrinos.

We compute our rates using the open source neutrino interaction
library \lstinline{NuLib}~\footnote{\url{http://www.nulib.org}}.
We compile a table of sources defined over the four dimensions of
density, temperature, electron-fraction, and neutrino energy, and
interpolate quad-linearly to the points sampled along each ray.

\begin{table}%[h]
  \caption{
    We analyze neutrino interaction processes in terms of these categories.
    $\nu$ without a label represents a neutrino or antineutrino of any flavor,
    $N$ represents a nucleon $n$ or $p$,
    ${}^ZA$ represents a nucleus with mass number $A$ and charge $Z$, and
    $\gamma$ represents a high-energy photon.
    A prime indicates a change in that particle's energy.
  }
  \label{tab:neutrino_processes}
  \begin{tabularx}{\columnwidth}{X X}
    \hline \hline
    absorption/emission
    & $\nu_e + n \leftrightarrow e^- + p$                          \\
    & $\bar{\nu}_e + p \leftrightarrow e^+ + n$                    \\
    & $\nu_e + {}^AZ \leftrightarrow e^- + {}^A(Z+1)$              \\
    & $\bar{\nu}_e + {}^AZ \leftrightarrow e^+ + {}^A(Z-1)$        \\
    \hline
    elastic scattering
    & $\nu + N \leftrightarrow \nu + N$                            \\
    & $\nu + {}^AZ \leftrightarrow \nu + {}^AZ$                    \\
    \hline
    inelastic scattering
    & $\nu + e^- \leftrightarrow \nu' + e^{-'}$                    \\
    & $\nu + e^+ \leftrightarrow \nu' + e^{+'}$                    \\
    \hline
    thermal pair processes
    & $\nu + \bar{\nu} \leftrightarrow e^{-} + e^{+}$              \\
    & $\nu + \bar{\nu} + N + N \leftrightarrow N' + N'$            \\
    & $\nu + \bar{\nu} \leftrightarrow \gamma$                     \\
    % & $\nu_\alpha + \bar{\nu}_\alpha \leftrightarrow \nu_\beta + \bar{\nu}_\beta $ \\
    \hline \hline
  \end{tabularx}
\end{table}

\subsection{Absorption and emission via charged current}
\label{ssec:sources_ae}
At neutrino and thermal energies well below the masses of the muon or tauon
($m_{\mu}\sim100\,{\rm MeV}$)
only charged current processes involving $\nu_e$ and $\bar{\nu}_e$ are allowed.
For each of these processes, $i$, we may write an emission and absorption
coefficient as a function of the interaction cross-section
(e.g. \cite[Eqn.~A5]{brue1985-core_collapse}):
\todo{explain Fermi-blocking}
\begin{align}
  \label{eqn:sources_ae_1}
  C_{{\rm AE},i}
  &= \varepsilon j_i (1-f) - \varepsilon \chi_{a,i} f \\
  \label{eqn:sources_ae_2}
  &= \varepsilon j_i - (\varepsilon j_i+\varepsilon \chi_{a,i}) f,
\end{align}
where $j$ is the emissivity and $\chi_a$ the absorption opacity.
Both $j$ and $\chi_a$ have dimension ${\rm length}^{-1}$ and represent the
number of neutrinos emitted or absorbed per length traveled.
In radiation transport formulations using specific intensities instead of
distribution functions, an emissivity $\eta$ having dimension
${\rm energy}\,{\rm length}^{-3}\,{\rm time}^{-1}\,{\rm energy}^{-1}\,{\rm steradian}^{-1}$
is more commonly used.
The two are related by $j=(2\pi)^3\eta/\varepsilon^3$.
Note that for brevity we have suppressed the energy-dependence of
the terms $j$, $\chi_a$, and the distribution functions.

In the special case of radiative equilibrium we know that the source
term vanishes: an equal number of neutrinos are emitted from
and absorbed by the matter for any length traversed.
We also know in this case that the neutrino distribution function must be
$f^{\rm eq}$, the equilibrium Fermi-Dirac distribution function of
Eqn.~\ref{eqn:feq}.
With these facts we can rerrange Eqn.~\ref{eqn:sources_ae_1} to give us
Kirchoff's Law:
\begin{align}
  \label{eqn:kirchoffs_law}
  j_i &= \frac{\chi_{a,i}}{1-f^{\rm eq}}f^{\rm eq}, \\
  \label{eqn:stimulated_opacity}
  &= \chi^*_{a,i}\,f^{\rm eq},
\end{align}
where in Eqn.~\ref{eqn:stimulated_opacity} we have introduced the opacity
corrected for stimulated absorption, $\chi^*_{a,i}$.

Using these expressions and computing the sum over stimulated opacities
$\chi^*_{a,i}=\sum_i \chi^*_{a,i}$,
the invariant emissivity and opacity for absorption/emission are
\begin{align}
  \label{eqn:ae_emissivity_summed_2}
  \mathscr{E}_{{\rm AE}} &= \varepsilon \, \chi^*_a \, f^{\rm eq}, \\
  \label{eqn:ae_opacity_summed_2}
  \mathscr{K}_{{\rm AE}} &= \varepsilon \, \chi^*_a.
\end{align}
These expressions are equivalent to those in
Eqns.~\ref{eqn:ae_emissivity_summed} and \ref{eqn:ae_opacity_summed}.

We use the above treatment for $\nu_e$ and $\bar{\nu}_e$ only;
the $\mu$ and $\tau$ neutrinos and antineutrinos do not participate in
charged current absorption/emission interactions at these temperatures
and energies.
However we do use an effective stimulated absorption opacity $\chi^*_a$
for the heavy-lepton neutrinos, computed by \lstinline{NuLib} as
described in \cite{ocon2015-gr1d_with_nu} which follows
\cite{brue1985-core_collapse,burr2006-neutrino_opacities}.
This is described in App.~\ref{ssec:sources_pp}.

\subsection{Elastic scattering}
\label{ssec:sources_se}
Neutrino scattering on particles of mass much greater than $\varepsilon$
(i.e. nucleons) is essentially iso-energetic. Following 
\cite[Eqn.~A8]{brue1985-core_collapse} or
\cite[Eqn.~4.20]{shib2011-truncated_moment} the collision term for the
$i$-th process takes the form
\begin{equation}
  \label{eqn:sources_se_1}
  C_{{\rm SE},i}(\ell_\alpha)
  = \frac{\varepsilon^3}{(2\pi)^3}
  \oint d\Omega' R_{{\rm SE},i}(\omega')
  \Big(f(\ell'_\beta) - f(\ell_\alpha)\Big)
\end{equation}
where $R_{\rm SE,i}(\omega')$ is the scattering kernel for the $i$-th process
from direction $\ell'_\beta$ to direction $\ell_\alpha$
having dimension ${\rm energy}^{-1}$,
and the cosine of the scattering angle is
$\omega' \equiv \psi^{\alpha\beta}\ell_\alpha\ell'_\beta$, with
$\psi^{\alpha\beta}$ the inverse of the spacetime metric.
Note that for brevity in Eqn.~\ref{eqn:sources_se_1} we have suppressed the
energy dependence of all of the terms.

It is customary to approximate the scattering kernel to linear order in
$\omega$ (as in~\cite[Eqn.~4.21]{shib2011-truncated_moment}):
\begin{equation}
  \label{eqn:se_kernel_expansion}
  R_{{\rm SE},i}(\varepsilon,\omega) \approx
  R_{{\rm SE},i}^0(\varepsilon) + \omega \, R_{{\rm SE},i}^1(\varepsilon).
\end{equation}
Using this expansion, and the moments defined in Eqns.~\ref{eqn:J}
and \ref{eqn:Ha} we may write Eqn.~\ref{eqn:sources_se_1}:
\begin{equation}
  \label{eqn:sources_se_2}
  C_{{\rm SE},i}
  = \varepsilon \chi_{s,i} \Phi_i - \varepsilon \chi_{s,i} f,
\end{equation}
where we have introduced the scattering opacity $\chi_{s,i}$
and the background scattering field $\Phi$:
\begin{align}
  \label{eqn:opacity_s}
  \chi_{s,i}(\varepsilon)
  &= \frac{4\pi\varepsilon^3}{(2\pi)^3}R_{{\rm SE},i}^0(\varepsilon), \\
  \label{eqn:background_phi}
  \Phi_i(\varepsilon,\ell_\alpha)
  &= \frac{(2\pi)^3}{\varepsilon^3}\frac{1}{4\pi}
  \left(J(\varepsilon)+
  \frac{\chi^{\rm tr}_{s,i}(\varepsilon)}{\chi_{s,i}(\varepsilon)}
  \ell_\alpha H^\alpha(\varepsilon)\right),
\end{align}
with $\chi^{\rm tr}_{s,i}$ the transport scattering opacity,
\todo{drop anisotropic term and defend}
\begin{equation}
  \label{eqn:opacity_s_tr}
  \chi^{\rm tr}_{s,i}(\varepsilon)
  = \frac{4\pi\varepsilon^3}{(2\pi)^3}R_{{\rm SE},i}^1(\varepsilon).
\end{equation}

Thus computing sums over the opacities
$\chi_s\equiv\sum_i\chi_{s,i}$ and
$\chi^{\rm tr}_s\equiv\sum_i\chi^{\rm tr}_{s,i}$,
the invariant emissivity and opacity for elastic scattering are
\begin{align}
  \label{eqn:se_emissivity_summed}
  \mathscr{E}_{\rm SE}
  &= \varepsilon \, \chi_s \, \Phi, \\
  \label{eqn:se_opacity_summed}
  \mathscr{K}_{\rm SE}
  &= \varepsilon \, \chi_s.
\end{align}

We have made the energy-dependence of the background scattering field
explicit in Eqn.~\ref{eqn:background_phi}.
When we compute $\Phi(\varepsilon)$ using $J(\varepsilon)$ and
$H^\alpha(\varepsilon)$ from a multi-group
M1 transport evolution, we call this the \emph{spectral} method.

However moment evolutions with multiple energy groups are still rare.
Most simulations employ a gray moment scheme,
evolving only the energy-integrated moments, $J$ and $H^\alpha$
and sometimes the number density $G$.
If such is the case we resort to the \emph{gray} method,
by reconstructing the energy-dependent source terms
from gray moments, assuming a diluted Fermi-Dirac spectrum.
We use the following procedure:
\begin{enumerate}
\item
  Interpolate the fluid temperature, $T$,
  equilibrium neutrino chemical potential, $\eta_\nu$,
  and evolved neutrino energy and number densities in the fluid frame,
  $J$ and $G$, from the simulation grid.
\item
  Compute the average neutrino energy in the fluid frame
  \begin{equation}
    \langle \varepsilon \rangle \equiv J/G.
  \end{equation}
\item
  Compute the neutrino spectral temperature, assuming equilibrium with the
  fluid
  \begin{equation}
    \label{eqn:spectral_temp}
    T_\nu = \langle \varepsilon \rangle
    \frac{\mathscr{F}_2(\eta_\nu)}{\mathscr{F}_3(\eta_\nu)},
  \end{equation}
  where $\mathscr{F}_b$ is the Fermi integral
  \begin{equation}
    \label{eqn:fermi_integral}
    \mathscr{F}_b(\eta) = \int_0^\infty dx \, x^b (1+e^{x-\eta})^{-1}.
  \end{equation}
\item
  Assume the background neutrino fields have the same total density
  as the evolved moments
  \begin{align}
    \label{eqn:J_from_gray}
    J(\varepsilon) &=
    J \frac{\varepsilon^3}{T_\nu^4 \,\mathscr{F}_3(\eta_\nu)}
    (1+\exp(\varepsilon/T_\nu-\eta_\nu)^{-1}, \\
    \label{eqn:H_from_gray}
    H^\mu(\varepsilon) &=
    H^\mu \frac{\varepsilon^3}{T_\nu^4\, \mathscr{F}_3(\eta_\nu)}
    (1+\exp(\varepsilon/T_\nu-\eta_\nu)^{-1}.
  \end{align}
\end{enumerate}
When we use these moments in Eqn.~\ref{eqn:background_phi},
we call this the \emph{gray} method.

\subsection{Inelastic scattering}
\label{ssec:sources_si}
Neutrino scattering off of electrons is inelastic, changing the
magnitude and direction of the species' momenta.
\todo{cite \cite{keil2003-pinched_spectra} or someone}
A similar formalism to the above could be used to derive source terms
for inelastic scattering.
\todo{compare relative amounts of inelastic to elastic scattering}
We save that for future work, pointing out here that inelastic scattering
treated this way is very sensitive to the energy-dependence of
the background field.

\subsection{Thermal pair annihilation and production}
\label{ssec:sources_pp}
We do not include thermal pair processes within the above formalism,
\todo{explain this choice, and cite \cite{keil2003-pinched_spectra} or someone}
but as mentioned in App.~\ref{ssec:sources_ae}, we do include
these processes in an effective absorption opacity for $\mu$ and $\tau$
neutrinos and antineutrinos. \lstinline{NuLib} follows
\cite{brue1985-core_collapse,burr2006-neutrino_opacities}
by computing the emissivity of the pair processes in
\todo{describe better}
Tab.~\ref{tab:neutrino_processes}, and applying Kirchoff's law,
Eqn.~\ref{eqn:kirchoffs_law}, to convert this to an opacity.
We only use this effective opacity for heavy-lepton neutrinos because
it is subdominant to the true absorption opacity for
$\nu_e$ and $\bar{\nu}_e$.

\subsubsection*{Acknowledgements}
This research is supported in part by the Joint Institute for Nuclear
Astrophysics---Center for the Evolution of the Elements (JINA-CEE)
and the National Science Fondation under grant numbers X. \todo{grant nos?}
We thank Sherwood Richers for providing his neutrino transport code
\lstinline{Sedonu} for comparisons,
Fran\c{c}ois H\'{e}bert for his timely help in finding and
correcting bugs in the time-interpolation algorithm,
and Luke Roberts for helpful discussions.

% ******************************************************************************
\bibliography{references}

\end{document}